### 2021-11-09 prism-backend/research/soa_review/discussion.tex
Added comment on CKS10.

\section{Discussion}
\label{sec:discussion}

Interesting topic for discussion, which applies mostly for ACs, but probably
to most other equivalent tools: we have tools that provide a (reasonably
complete and) secure way to authenticate in a privacy-preserving manner.
However, what do we do after such authentication takes place? E.g., it is of
little use to employ an AC to log into an online service, if that service
will learn all about us after we have anonymously authenticated. How does
privacy-preserving authentication improve the situation even in that case?
Is there some way in which we can help avoid that?

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "sok-privsig"
%%% End: