### 2021-11-09 prism-backend/research/soa_review/soa-review.tex

\documentclass{llncs}%[10pt]{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{hyperref}

\title{State of the Art Review on Privacy-Preserving Signatures, Credentials, and Authentication}
\author{}

% Custom shorthands for this paper
\input{symbols.tex}

\begin{document}
\maketitle

\begin{abstract}
	\textbf{DISCLAIMER}: Please, keep in mind that, as of now, this document
	basically contains a research log in the shape of personal notes. Do not take 
	anything written here as ground truth.
\end{abstract}

\input{introduction.tex}
\input{gs.tex}
\input{abs.tex}
\input{ac.tex}
\input{daa.tex}
\input{other.tex}
\input{discussion.tex}
\input{conclusion.tex}

\section{TODOs}

General/pending doubts and tasks:

\begin{itemize}
\item Many constructions of ACs and GSs seem to be built using the same
  approach: an issuer and a user interact so that the latter receives a signature
  from the former; then the user employs that signature to do something else (in
  ACs, prove attributes; in GSs, produce further signatures). Still, the typical
  security models are different. Concretely, GSs have a very well established
  model, but ACs don't. It also ``bothers'' me that none of the AC models I've
  seen cover something similar to non-frameability, nor traceability, but just
  include conventional unforgeability properties. Why is this?
\end{itemize}

\bibliographystyle{splncs04}
\bibliography{soa-review}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: ucl.tex
%%% End: