### 2021-11-29 prism-backend/research/soa_review/soa-review.tex

\documentclass{llncs}%[10pt]{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{hyperref}

\title{State of the Art Review on Privacy-Preserving Authentication}
\author{}

% Custom shorthands for this paper
\input{symbols.tex}

\begin{document}
\maketitle

\begin{abstract}
  \comment{\textbf{DISCLAIMER}: Please, keep in mind that this is a working
    draft. Its main use in its current state is to foster debate on the topics
    it deals with. Do not take anything written here as ground truth.}

  This document explores three closely related cryptographic systems in the area
  of privacy-enhanced and, more concretely, in the domain of authentication.
  Namely, we study advances in group signatures (GS), anonymous credentials
  (AC), and direct anonymous attestation (DAA). All three have somehow similar
  purpose, have followed similar approaches, and have received much attention in
  the the academic literature for the last (almost) four decades but, in some
  sense, independently on one another.

  Our goal is to extract common insight from all of them. Firstly, from a
  theoretical perspective, with the aim to identify the advantages and
  disadvantages of existing models and approaches, and develop intuition as to
  whether we can (or cannot) combine ideas and fill gaps, and what may be the
  most suitable path to follow for new projects, depending on their needs.
  Secondly, from a practical perspective, DAA and ACs have already been deployed
  in the real world, and analysing the difficulties they have come across will
  certainly be of use.

\end{abstract}

\iffalse
\input{introduction.tex}
\fi
\input{gs.tex}
%\input{abs.tex}
\input{ac.tex}
\input{daa.tex}
\input{other.tex}
\input{discussion.tex}
\input{conclusion.tex}

\bibliographystyle{splncs04}
\bibliography{soa-review}

\appendix

\section{TODOs}

General/pending doubts and tasks:

\begin{itemize}
\item Many constructions of ACs and GSs seem to be built using the same
  approach: an issuer and a user interact so that the latter receives a signature
  from the former; then the user employs that signature to do something else (in
  ACs, prove attributes; in GSs, produce further signatures). Still, the typical
  security models are different. Concretely, GSs have a very well established
  model, but ACs don't. It also ``bothers'' me that none of the AC models I've
  seen cover something similar to non-frameability, nor traceability, but just
  include conventional unforgeability properties. Why is this?
  \begin{itemize}
  \item Concerning non-frameability, this probably is because the de-anonymizing
    functionality that ACs include (if any) is revocation, which is essentially
    a blacklisting. On the other hand, GSs usually allow opening, which extracts
    the actual identity. Opening is ``stronger'' than blacklisting, in the sense
    that it can be used to build blacklisting, but it is also more threatening
    to privacy and may have further implications. Among these, unequivocable
    framing of users. This, in general, does not seem possible with conventional
    blacklisting (or, at least, not in an unequivocable way).
  \item Concerning traceability. \cite{fhs19} definition of unforgeability
    resembles traceability. I guess this would be the case in most AC works.
  \end{itemize}
\end{itemize}

\input{app-gs}
\input{app-ac}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: ucl.tex
%%% End: