### 2021-12-13 prism-backend/research/single-mnemonic/discussion.tex

\section{Discussion}
\label{sec:discussion}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "single-mnemonic"
%%% End: