### 2021-12-13 prism-backend/research/single-mnemonic/introduction.tex

\section{Introduction}
\label{sec:introduction}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "single-mnemonic"
%%% End: