### 2021-12-13 prism-backend/research/single-mnemonic/single-mnemonic.tex

\documentclass{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
%\usepackage{amssymb}
\usepackage{xspace}
%\usepackage[normalem]{ulem}
\usepackage{hyperref}

\title{Research Spike: Evaluation of Using a Single Mnemonic both
  for Cryptocurrency and Identity Wallets}
\author{}

% Custom shorthands for this paper
\input{symbols.tex}

\begin{document}
\maketitle

\begin{abstract}
  \comment{\textbf{DISCLAIMER}: Please, keep in mind that this is a working
    draft. Its main use in its current state is to foster debate on the topics
    it deals with. Do not take anything written here as ground truth.}  
\end{abstract}

\input{introduction.tex}
\input{bip32-etal.tex}
\input{idwal.tex}
\input{discussion.tex}
\input{conclusion.tex}

\bibliographystyle{splncs04}
\bibliography{single-mnemonic}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: ucl.tex
%%% End: