### 2021-12-18 prism-backend/research/single-mnemonic/iowal.tex

\section{IO Ecosystem Specifics}
\label{sec:io}

\paragraph{Lightwallet.} HD wallets in Cardano are mostly BIP44-compatible
wallets, with the exception that the used elliptic curve is Ed25519,
instead of Secp256k1, like Bitcoin's BIP32 (and thus, BIP44). This
difference is marked by using index 1852 for hardened child derivation at
layer 1 (purpose level), as specified in CIP1852%
\footnote{\url{https://cips.cardano.org/cips/cip1852/}. Last access, December
  15th, 2021.}.

\paragraph{Atala.} HD wallets in Atala follow a slightly different tree
organisation, layer-wise. Namely, it defines 4 layers: root, DID number, key
type, and index. All child derivations are done in hardened mode. The elliptic
curve is Secp256k1, as BIP44. See the key derivation document for details%
\footnote{\url{https://github.com/input-output-hk/atala-prism/blob/master/prism-backend/docs/protocol/key-derivation.md}. Last access December 15th, 2021. \textbf{WARNING!
    Internal link!}}.

\paragraph{Related effots.} There seems to be an effort to analyse the security
of wallets that manage addresses combining payment and staking keys \cite{kkl20}.
These wallets are referred to as PoS wallets. The paper identifies a series of
malleability attacks, depending on how the addresses are derived from the
associated staking and payment keys. A core wallet functionality is proposed,
and a base protocol realizing that functionality is given. While the proposal
seems to be somehow compatible with BIP44 wallets, the proposal seems to be
somehow orthogonal to the way keys are derived -- instead, they focus on how
to generate addresses, and check that produced/received addresses are correct.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "single-mnemonic"
%%% End: