### 2022-01-10 prism-backend/research/single-mnemonic/bip32-etal.tex
Changes based on Tony's comments.

\section{An Overview of BIP32-based HD Wallets}
\label{sec:bip32}

In this section, we give a brief overview of the BIP32 specification, and
provide a ``full path'' from mnemonic to key pair. To avoid repeating already
well documented processes, the overview will refer to the specifications when
appropriate.
%
In a final subsection, we will also overview how does all this apply to the
Cardano Lightwallet and Atala Prism cases.

\subsection{Main Specifications: A Brief History of BIPs for Wallets}

BIPs, or Bitcoin Improvement Proposals, are ``standards'' in the Bitcoin domain
that specify how concrete components of the ecosystem work and/or should be
implemented. Being Bitcoin the first blockchain, many others import these
standards.

In 2013, BIP32 was published. It defines an approach to HD wallets
using a tree structure in which each node can be associated to a key pair. It
also defines two ways to derive child nodes from parent nodes: one that allows,
given a public key of a node, to derive the public and private keys of all its
descendants; and another one that requires the private key for derivation of
its child nodes. Based on this child derivation processes, it also specifies a
basic strategy for building HD structures for payments.

However, although BIP32 does specify a basic structure, it does not mandate that
it has to be followed. According to BIP43%
\footnote{\url{https://github.com/bitcoin/bips/blob/master/bip-0043.mediawiki}.
  Last access, December 13th, 2021.}, published in 2014, this led to a set of
self-claimed BIP32-compliant wallets not to be compliant between themselves.
BIP43 requires that the first level of child derivation (directly from the
root) be dedicated to define a ``purpose''. BIP44%
\footnote{\url{https://github.com/bitcoin/bips/blob/master/bip-0044.mediawiki}.
  Last access, December 13th, 2021.}, published simultaneously to BIP43, further
refines this, by establishing 6 different levels (which we introduce in
subsequent sections). Thus, BIP44 is an specialization of BIP43, which is itself
an specialization of BIP32.

Somehow orthoghonally to these, BIP39 defines how to encode random binary seeds
into mnemonic sentences, and back. In this manner, the root node of a BIP44
wallet is decoded into a bitstring, from a mnemonic, using BIP39. Then, this
root node and its child nodes are computed as defined in BIP44. \tabref{tab:bips}
summarizes the mentioned standards.

\begin{table}[ht!]
  \centering
  \scalebox{0.95}{
    \begin{tabular}{|c|c|l|}
      \hline
      \bf BIP & \bf Year & \bf Description \\
      \hline
      BIP32 & 2013 & Basic structure and derivation rules for HD wallets. \\
      BIP43 & 2014 & Definition of ``purpose'' level. \\
      BIP44 & 2014 & Definition of 6-level structure (root, purpose, coin type,
                     account, change, address.) \\
      BIP32 & 2013 & Encoding/decoding of bitstrings to/from mnemonic sentences. \\
      \hline
    \end{tabular}
  }
  \caption{Summary of BIPs for HD wallets.}
  \label{tab:bips}
\end{table}

HD wallets that follow the BIP44 specification have to define a value of 44
(with hardened derivation; see next) at the ``purpose'' layer. Wallets following
a different structure, but still compatible with BIP32, can (should?) define
their own code, for the sake of ensuring compatibility across implementations.
For instance, Cardano defines the code 1852 (again, hardened derivation), as
it has some minor variations with respect to BIP44. There are several well-known
wallets that follow this specification. For instance, Trezor%
\footnote{\url{https://docs.trezor.io/trezor-firmware/misc/coins-bip44-paths.html}.
  Last access, January 10th, 2022.}, or Ledger%
\footnote{\url{https://support.ledger.com/hc/en-us/articles/4404388633489-Export-your-accounts?docs=true}.
Last access, January 10th, 2022.}.

Additionally, Satoshi Labs SLIP44 maintains a registry of the different
cryptocurrencies that have ``registered'' codes for wallet specifications
that follow BIP44. That is, this registry contains codes for the ``coin type''
level of BIP44. Examples of registered coins include Bitcoin mainnet and testnet,
Litecoin, Cardano, Ether, and many others (currently, there are several
hundreds).

% While the main specification is BIP32, it essentially gives the basic guidelines
% to create HD wallets. To avoid ambiguity leading to incompatible
% implementations, BIP43 and BIP44 specify how to organize HD wallets so
% that the resulting tree structure enables the most common use cases (which, at
% least initially, focus on the cryptocurrency scenario). In addition, the
% specification that defines how to encode (pseudo-)random bitstrings as
% mnemonic is given in BIP39.

\subsection{Main Concepts}
\label{ssec:mainhdconcepts}

\begin{description}
\item[HD wallet tree structure.] HD wallets are organised as trees, where the
  root node is directly derived from the main seed (typically, a mnemonic-encoded
  random string). We denote the root level with level $0$, and all descendants
  with increasing numbers. A level other than the root level can have an
  arbitrary number of nodes.
\item[(Child) Index.] The child index just identifies how many childs can be (or
  have been) created before the current child, at the current depth of the tree.
  We assume $0$ to be the first index.
\item[Chaincode.] To introduce extra ``non-determinism'' in the way child keys
  are derived from parent keys, part of the pseudo-randomly derived data is
  used as key for pseudo-random derivations of lower levels. This part of the
  derived data is referred to as chaincode. We will denote the chaincode of the
  $j$-th node at level $i$ of the tree with $c_{i,j}$. This data should be kept
  secret; otherwise, entire subtrees can be compromised.
\item[Extended keys.] An extended public (resp. private) key is composed by the
  public (resp. private) key and the chaincode.  
\item[Hardened child keys.] A hardened child key can only be derived from the
  parent private key, and the parent key chaincode. 
\item[Non-hardened child keys.] A non-hardened child key can be derived both
  from the parent private and parent public key, and the parent key chaincode.
\end{description}

\paragraph{Hardened vs non-hardened child derivation.} Basically, child nodes
are derived as rerandomizations of their parent node private key. The
pseudo-random data used for rerandomization can be derived from the parent
private extended key, or from the parent public extended key. The former case is
referred to as hardened derivation, and the latter as non-hardened derivation.
In a nutshell, this means that hardened child keys (private or public) can only
be derived from their parent private key, but not from their parent public key.
Naturally, child private keys can only be derived from parent private keys.

For self-containedness, \figref{fig:childderivation} shows the different
algorithms to derive child keys from a parent keys, excluding some details. For
the full algorithms, we refer to BIP32.

\begin{figure}[ht!]
  \begin{minipage}[t]{0.55\textwidth}
    \procedure{$prv2HardChild$ \pccomment{Parent private key $\rightarrow$ $i$-th Child hardened key pair}}{
      \pcind I \gets HMAC(c_{par}, k_{par}, i) \\
      \pcind k^h \gets (k_{par} + Left(I)) \bmod n; K^h \gets k^hG; c^h \gets Right(I) \\
      \pcind \pcreturn (k^h,K^h,c^h) \\
    }
    \procedure{$prv2NonHardChild$ \pccomment{Parent private key $\rightarrow$ $i$-th Child non-hardened key pair}}{    
      \pcind I \gets HMAC(c_{par}, K_{par}, i) \\
      \pcind k^{nh} \gets (k_{par} + Left(I)) \bmod n; K^{nh} \gets k^{nh}G; c^{nh} \gets Right(I) \\
      \pcind \pcreturn (k^{nh},K^{nh},c^{nh}) \\      
    }
    \procedure{$pub2NonhardPubChild$ \pccomment{Parent public key $\rightarrow$ $i$-th Child non-hardened public key}}{      
      \pcind I \gets HMAC(c_{par}, K_{par}, i) \\
      \pcind K^{nh} \gets Left(I)G + K_{par}; c^{nh} \gets Right(I) \\
      \pcind \pcreturn (K^{nh},c^{nh}) 
    }    
  \end{minipage}
  \caption{Algorithms for child derivation. $k_{par},K_{par},c_{par}$ denote the
    parent's public key, private key, and chaincode, respectively. $k,K,c$
    denote the resulting child's private key, public key, and chaincode. $h$ or
    $nh$ superscripts denote hardened or non-hardened derivation. $Left$
    (resp. $Right$) divides a bitstring in two equal parts, and takes the left
    (resp. right) half. The result of both $Left$ and $Right$ can be interpreted
    as a scalar, and thus multipled by points in the elliptic curve, such as $G$,
    which is the base point, or public keys, which also are points in the
    curve.}
  \label{fig:childderivation}
\end{figure}

\paragraph{Isolation of HD wallet subtrees.} In the following section, we use
the concept of isolated and non-isolated subtrees. This is something that is
only indirectly mentioned in the BIPs and other related works, but which is of
crucial importance in order to limit the impact of compromised keys. Namely,
assume that a child key is derived in non-hardened mode. Then, if the extended
parent public key is leaked, as well as the private (non-extended) non-hardened
child key, it is possible to extract the parent private key -- and, therefore,
all the keys (private and public; hardened or not) that derive from it. If child
keys are derived in hardened mode, this does not apply. For the sake of clarity,
we illustrate the difference through the diagram in \figref{fig:isolation}.

\begin{figure}[ht!]
  \includegraphics[width=\textwidth]{figures/child_derivation.png}
  \caption{Sub-tree isolation through hardened child derivation. $G$ is the
    base point of the underlying elliptic curve, and $n$ is the order of
    the associated group.}
  \label{fig:isolation}
\end{figure}

For instance, assume that the extended public key $(K_{i,j}, c_{i,j})$ is
compromised, along with its child private non-hardened key $k^{nh}_{(i+1),t}$.
Then, it is straightforward  to recompute the \emph{parent} private key
$k_{i,j}$ (independently on whether it was derived in hardened mode or not) as:
$k_{i,j} \leftarrow (K_{(i+1),t}^{nh} - Left(I_{(i+1),t}^{nh})) \bmod n$.
Obviously, given the parent private key $k_{i,j}$, it is straightforward to
derive all its descendants (hardened or not), as well as possibly and
recursively apply the same strategy, if the just recovered parent key was
derived in non-hardened mode.

\subsection{From Mnemonic to Key Pairs: High-level Overview}

The following description is based on BIP39 for the processing of mnemonics,
and BIP44 (which is a restriction of BIP43 and BIP32) for general structure of
an HD wallet, which we depict in \figref{fig:bip44-tree}.

\paragraph{Path notation.} As can be seen in \figref{fig:bip44-tree}, the
hierarchical structure of an HD wallet is defined in layers (or levels), where
different indexes are used to derive the cryptographic values associated to
different nodes in a same layer, with the exception of the root node (at layer
0), which is frequently just denoted with an \texttt{m}, after master seed.
Thus, for instance, paths  \texttt{m/1} and \texttt{m/2} denote a tree with two
layer 1 nodes computed with index \texttt{1} and index \texttt{2},
respectively. If a child node index is \emph{primed}, it means that hardened
derivation is used (whereas the absence of a prime means non-hardened
derivation). More specifically, since BIP44 wallets have up to layer 6, paths
look like \texttt{m/a'/b'/c'/d'/e/f}. Note that layers 1 to 3 are derived in
hardened mode, and layers 4 and 5 are derived in non-hardened mode.

\begin{figure}[ht!]
  \includegraphics[width=\textwidth]{figures/bip44_tree.png}
  \caption{Basic tree structure of a BIP44-based HD wallet.}
  \label{fig:bip44-tree}
\end{figure}

\paragraph{From (pseudo-)random bitstrings to mnemonics, and back.} %
Mnemonics encode randomness in human-readable form. BIP39 is the main
specification for this. Roughly, between 128 and 256 random bits (in multiples
of 32) are generated, a checksum (the first few bits of a hash of these random
bits) is concatenated, and the result is divided in chunks of 11 bits. Each of
these chunks is used as index to a predefined dictionary of words. The result
is a mnemonic of 12 to 24 words. To convert from mnemonic to random bits, just
the inverse process has to be followed (verifying the checksum).

Given the mnemonic, it is converted back into a bitstring as just described.
Then, this bitstring is processed with an HMAC function, using the string
``Bitcoin seed'' as key based on SHA512 -- hence producing 512 bits of
output. The first 256 bits are set as master secret key, and the last 256 bits
as master chaincode (i.e., $k_{0,0}$ and $c_{0,0}$, in our previous notation).

\paragraph{Layer 1: \emph{isolated} subtrees per \texttt{purpose}} %
The first layer of a BIP44-compatible HD wallet is composed of up to $2^{31}$
nodes of hardened childs: i.e., layer 1 is composed of childs $(1,2^{31})$
to $(1,2^{62}-1)$, where the second element in the pair is the index $i$,
whose value is used in the child derivation processes, and denotes for what
``purpose'' the descendant keys will be used -- this roughly translates to the
system (e.g., Bitcoin, Cardano... or even non-blockchain systems, in theory).

Note that, being layer 1 computed in hardened mode, a private key compromise
of one node in layer 1 does not affect the master node, nor other layer 1
siblings. Hence, this is an ``isolated'' subtree, according to our terminology.

For BIP44-compatible wallets, the expected value for the $i$ index is $44$, so
all paths begin with \texttt{m/44'}. Cardano, since the Shelley era, uses index
$1852$ (and thus, Cardano paths begin with \texttt{m/1852'}%
\footnote{\url{https://cips.cardano.org/cips/cip1852/}. Last access, December
  14th, 2021.}. Note that, since these are hardened indexes, the encoding
rules translate this into $1852+2^{31}$ (resp. $44+2^{31}$ for BIP44); see
BIP44 for more details.

\paragraph{Layer 2: \emph{isolated} subtrees per \texttt{coin type}} %
The second layer is again composed of up to $2^{31}$ nodes of hardened childs:
i.e., layer 2 is composed of childs $(2,2^{31})$ to $(2,2^{62}-1)$ per each child
node of layer 1. In theory, up to $2^{31}*2^{31}=2^{62}$ second-level nodes can
coexist in the same HD wallet. The aim of the second level is for any HD wallet
to be able to support multiple types of coins. E.g., so that the same mnemonic
can be used to derive key pairs for Bitcoin, Cardano, etc.

Note that, being layer 2 computed in hardened mode, a private key compromise
of one node in layer 2 does not affect its parent in layer 1, nor other layer 2
siblings. Hence, this is an ``isolated'' subtree, according to our terminology.

Although not mandatory, each coin is expected to have an assigned coin type
index. The complete list of reserved indexes is available in SLIP44%
\footnote{\url{https://github.com/satoshilabs/slips/blob/master/slip-0044.md}.
  Last access, December 14th, 2021.}. For instance, Cardano's ADA has an index
of 1815 in hardened mode -- thus, its path prefix is \texttt{m/1852'/1815'}.

\paragraph{Layer 3: \emph{isolated} subtrees per \texttt{account}} %
The third layer is again composed of up to $2^{31}$ nodes of hardened childs:
i.e., layer 3 is composed of childs $(3,2^{31})$ to $(3,2^{62}-1)$ per each child
node of layer 2. In theory, up to $2^{3*31}=2^{93}$ third-level nodes can
coexist in the same HD wallet. The aim of the third level is to allow the HD
wallet user to maintain many accounts per coin type. E.g., the same mnemonic
seed could be used for up to $2^{31}$ key pairs for Cardano addresses.

Again, note that being layer 3 computed in hardened mode, a private key compromise
of one node in layer 3 does not affect its parent in layer 2, nor other layer 3
siblings. Hence, this is an ``isolated'' subtree, according to our terminology.

\paragraph{Layer 4: \emph{non-isolated} subtrees per \texttt{change}}
According to BIP32, the fourth layer can be composed of up to  $2^{31}$ nodes of
non-hardened childs: i.e., layer 4 can be composed of childs $(4,0)$ to
$(4,2^{31}-1)$ per each child node of layer 3. However, in BIP44, only
non-hardened childs 0 and 1 are used: i.e., each node of layer 3 will only have
non-hardened childs $(4,0)$ and $(4,1)$. This layer is used to differentiate
whether the descendant nodes will be used as change addresses, or not. I.e.,
child $(4,0)$ will \emph{not} be used for receiving change in payments, and
therefore is expected to be associated to an address that may be published
outside of the wallet; while child $(4,1)$ is expected to be used for receiving
change in payments, and will typically not be published outside of the
transaction.

Note that, since nodes in layer 4 are not derived in hardened mode, if the private
key of a layer 4 node is compromised, along with the extended public key of its
layer 3 parent, then the entire subtree of that layer 3 node will be compromised
(but not the subtrees of other layer 3 nodes, as layer 3 is derived in hardened
mode). For example, take the derivation path of a layer 4 node to be
\texttt{m/x'/y'/z'/0}. If the private key of \texttt{m/x'/y'/z'/0} is leaked to
an attacker, who also gains access to the \emph{extended public} key (i.e.,
the public key and chaincode) of \texttt{m/x'/y'/z'}, then the attacker will
be able to compute the extended private (and public) keys of the sibling node
\texttt{m/x'/y'/z'/1}. This is precisely the attack explained in the last
paragraph of \secref{ssec:mainhdconcepts}.

\paragraph{Layer 5: \emph{non-isolated} final key pairs (\texttt{index})} %
The fifth and last layer is used to derive final key pairs. It can be composed
of up to $2^{31}$ non-hardened nodes: i.e., every change (and non-change) node
of level 4 may have level 5 childs from $(5,0)$ to $(5,2^{31}-1)$.

Note that, since layer 4 nodes are derived in non-hardened mode, if a layer 5
private key is compromised, as well as its parent extended public key of layer
4, then the whole layer 4 subtree will be compromised. Furthermore, if the
layer 3 extended public key from which the layer 4 parent derives is also
compromised, the entire account will be compromised too, as layer 4 is also
derived in non-hardened mode.

\subsection{Comments on a Recent Security Evaluation of BIP32 Wallets}

Very recently, \cite{def+21} was published in CCS'21, a major applied
cryptography and security conference. The paper builds on prior work from
2019 \cite{dfl19}, and analyses the security of BIP32-like wallets. For this purpose,
the authors first model and analyse (additively) re-randomizable signatures
built on ECDSA. They prove security, as long as every message is signed at
most once, at the cost of a security loss proportional to the number of
signature re-randomizations performed by the adversary\footnote{They also prove
  this security loss to be unavoidable with additive re-randomization, although
  I have not checked that proof.}. Then, they build a model for HD wallets,
and provide a generic construction using additively re-randomizable signatures.
The properties considered in this HD wallet model are unlinkability and
unforgeability (restricted to only one signature per message). The unlinkability
property expects that an adversary cannot distinguish (uncompromised) keys
derived from a non-hardened node, from keys derived from keys derived from an
independent master key. The unforgeability property is the usual one, excluding
the one message restriction (which seems reasonable, as long as new randomness
is included in every signed message); namely, the adversary should not be able
to create a valid signature under a hardened key that has not been compromised,
or a non-hardened key.

An interesting extension is mentioned, to achieve forward unlinkability in case
of compromise of non-hardened node public key and chaincode. Namely, if that
happens, the adversary will be able to derive the public extended keys for
all the subtree, which breaks unlinkability for the subtree. The proposal is to
keep a state along with every node. On every child derivation in the tree, the
state of the existing nodes is refreshed. Then, in case of compromise of a
node, the keys previously derived in subtrees of it are not affected (assuming
that previous states are securely erased).

Building on the proven security of the additively re-randomizable ECDSA, they
prove security of their generic construction, which now also depends on the
number of leaked hardened keys. Assuming that roughly $1\%$ of the total number
of (additively re-randomized) hardened keys are compromised, the estimated
security level for ECDSA of 256-bit keys (hence, 128-bit security) is 91-bit
security.

\paragraph{Discussion.} %
The model does seem to follow BIP32, but \textbf{assuming a hot/cold approach}.
Concretely, they assume that non-hardened secret keys are always kept in a cold
wallet, and therefore cannot be compromised. On the other hand, hardened secret
keys are assumed \emph{not} to be stored in cold wallets, and therefore can be
compromised (this is reflected in the model with corresponding oracles or lack
thereof). The underlying reasoning being that hardened derivation is used to
share keys with not fully trusted entities (e.g., employees in a company). Also,
hardened keys are assumed to be leafs (i.e., keys used directly for signing); or,
alternatively, roots of independent subtrees. Under this modelling, if the
partially trusted entity leaks the hardened key, this would only affect the
security of its subtree (if it exists), but certainly not siblings or ancestors.
However, this may not be fit for BIP44 wallets, where hardened derivation is
expected to be used directly for the first 3 layers, and the leaves are actually
computed with non-hardened derivation.

The property of forward unlinkability is interesting for identity-related
wallets.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "single-mnemonic"
%%% End: