### 2022-01-18 prism-backend/research/single-mnemonic/appendix.tex
Final pass based on Tony's feedback.

\section{Appendix}

\subsection{Feedback Welcome}

Are you aware of further challenges that may be worth looking into? Have some
feedback? Please, reach out to us:

\begin{itemize}
\item Jesus Diaz Vico (Atala Semantics team).
\item Christos KK Loverdos (Atala Technical Director).
\item Ezequiel Postan (Atala Semantics team).
\item Tony Rose (Atala Head of Product).
\end{itemize}

\subsection{Further related work}

There seems to be a growing body of related research, which is specially
growing lately (probably, due to the rise of cryptocurrencies). The following
just includes some of the main references for self-bookkeeping (besides
\cite{kkl20} and \cite{def+21}, already mentioned in earlier sections). Note
that they, in turn, include references to further related work.

\begin{itemize}
\item ``Arcula: A Secure Hierarchical Deterministic Wallet for Multi-asset
  Blockchains'', \cite{lfa20}.
\item ``Simple, Efficient and Strongly KI-Secure Hierarchical Key Assignment
  Schemes'', \cite{fpp13}.
\item ``A Formal Treatment of Deterministic Wallets'', \cite{dfl19}.
\end{itemize}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "single-mnemonic"
%%% End: