### 2022-02-06 prism-backend/research/gsac/conclusion.tex
Creating branch for work on GSAC schemes.

\section{Conclusion}
\label{sec:conclusion}

\comment{This section is just a placeholder for now...}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "sok-privsig"
%%% End: