### 2022-02-06 prism-backend/research/gsac/preliminaries.tex

\section{Preliminaries}
\label{sec:preliminaries}

\subsection{Related Work}
\label{ssec:related}

\todo{\textbf{IMPORTANT!} There actually seems to be closely related work
  on Attribute-Based Group Signatures (ABGS): \cite{khad07,emo09,aa14}.}

\subsection{Building Blocks}
\label{ssec:bblocks}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "gsac"
%%% End: