### 2022-02-08 prism-backend/research/gsac/preliminaries.tex

\section{Preliminaries}
\label{sec:preliminaries}

\subsection{Related Work}
\label{ssec:related}

\todo{\textbf{IMPORTANT!} There actually seems to be closely related work
  on Attribute-Based Group Signatures (ABGS): \cite{khad07a,khad07b,emo09,aa14}.}

One difference between our current approach to GSAC, and the existing ABGS
works, is that in ABGS each group member can only get one credential. On the
other hand, we are trying to let any group member obtain as many credentials
as he wishes, possibly with different attributes, albeit all of them related to
the same user key pair (if a different user key pair is employed, then this is
considered as a different user).

\subsection{Building Blocks}
\label{ssec:bblocks}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "gsac"
%%% End: