### 2022-02-28 prism-backend/research/gsac/uas.tex

\documentclass{llncs}%[10pt]{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{hyperref}

\title{Universal Anonymous Signatures}
\author{}

% Custom shorthands for this paper
\input{symbols.tex}
\input{symbols-uas.tex}

\begin{document}
\maketitle

\begin{abstract}
  \comment{\textbf{DISCLAIMER}: Please, keep in mind that this is a working
    draft. Its main use in its current state is to foster debate on the topics
    it deals with. Do not take anything written here as ground truth.}  
\end{abstract}

\input{introduction-uas.tex}
\input{preliminaries-uas.tex}
\input{model-uas.tex}
\input{construction.tex}
%\input{benchmarks.tex}
\input{conclusion-uas.tex}

\bibliographystyle{splncs04}
\bibliography{uas}

%\appendix

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End: