### 2022-04-02 prism-backend/research/gsac/conclusion.tex
Merged GSAC and UAS into uas.tex paper. Defining concrete instantiation of GSAC.

\section{Conclusion}
\label{sec:conclusion}

\comment{This section is just a placeholder for now...}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End: