### 2022-04-04 prism-backend/research/gsac/sec-gsac.tex

\section{\GSAC: Merging Group Signatures and Anonymous Credentials}
\label{sec:gsac}

\todo{Some intro here}

\input{model-gsac.tex}
\input{construction-gsac.tex}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End: