### 2022-04-04 prism-backend/research/gsac/sec-uas.tex

\section{From \GSAC to \UAS}
\label{sec:uas}

\todo{Some intro here}

\input{model-uas.tex}
\input{construction-uas.tex}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End: