### 2022-05-02 prism-backend/research/gsac/security-gsac.tex

\subsection{Correctness and Security of \GSACGen}
\label{ssec:security-gsac}

For the sake of brevity, here we just enuntiate the main theorems related to
correctness and security of \GSAC, and defer the full proofs to
\appref{sapp:gsac-proofs}.

\begin{theorem}[Correctness of \GSACGen]
  \label{thm:correctness-gsac}
  If the underlying schemes for \todo{xxx}, our generic construction \GSACGen
  satisfies correctness as defined in \defref{def:correctness-gsac}.
\end{theorem}

\begin{theorem}[Anonymity of \GSACGen]
  \label{thm:anonymity-gsac}
  If the underlying $\NIZK^{\NIZKRel_{\Sign}}$ scheme is zero-knowledge and
  simulation-extractable, and the encryption system is IND-CCA secure,
  our \GSACGen construction satisfies anonymity as defined in
  \defref{def:anonymity-gsac}.
\end{theorem}

\begin{theorem}[Traceability of \GSACGen]
  \label{thm:trace-gsac}
  If the underlying $\NIZK^{\NIZKRel_{\Sign}}$ scheme is sound, the
  $\NIZK^{\NIZKRel_{\Open}}$ is  complete, the public-key encryption algorithm
  is correct, and the signature scheme for blocks of committed messages is
  one-more unforgeable, then our \GSACGen construction satisfies traceability as
  defined in \defref{def:trace-gsac}, except with negligible
  probability.
\end{theorem}

\begin{theorem}[Non-frameability of \GSACGen]
  \label{thm:frame-gsac}
  If the underlying scheme for $\NIZK^{\NIZKRel_{\Issue}}$,
  $\NIZK^{\NIZKRel_{\Sign}}$ and $\NIZK^{\NIZKRel_{\Open}}$ are sound,
  zero-knowledge, and simulation-extractable, and
  the commitment scheme \C is hiding, then our generic \GSACGen construction
  satisfies non-frameability as defined in \defref{def:frame-gsac}, except with
  negligible probability.
\end{theorem}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "gsac"
%%% End: