### 2022-05-03 prism-backend/research/gsac/sec-gsac.tex

\section{\GSAC: Merging Group Signatures and Anonymous
  Credentials}
\label{sec:gsac}

\todo{I think we should extend ``vanilla'' \GSAC to return the full \Attrs
  in \Open, but without needing \trans as a parameter to \Open (for compatibility
  with \UAS). This can be done by encrypting all attributes, including their
  ciphertext in the signature, and proving they are part of the credential. This
  would render the scheme too costly, but helps in arguing in favour of \UAS.
  Still, leaving that for after having updated the \UAS section, to see how
  it fits the global picture.}

Before moving into our full-fledged scheme for universal anonymous signatures,
we take a first step which, albeit somehow obvious and simple, helps understand
our final goal. Namely, we give a scheme that combines in a quite direct manner
the functionality of group signature schemes, and that of anonymous credential
schemes.

\paragraph{Why is this merge necessary in the real world?} %
Despite being an \emph{a priori} simple combination of both schemes we argue
that such a combination:

\begin{itemize}
\item Enables a natural way to incorporate accountability within 
  privacy-preserving identities. Furthermore, it does so in a formal way,
  via the traceability notion and, crucially, the non-frameability one,
  (which is not present in anonymous credentials).
\item Provides meaning to the conventional \Open functionality of group
  signatures, which traditionally has only required that an abstract concept of
  ``identity'' be returned. Concretely, assuming honest issuance (which is
  already a requirement in anonymous credentials; and in group signatures for
  traceability), we can make specific claims about the attributes associated to
  de-anonymized signers. Otherwise, without integrating attributes within the
  issued credentials, the value returned by \Open serves barely for blacklisting
  or similar functionalities -- but otherwise gives no information straight away
  about the signer's identity, forcing implementors to rely to non-cryptographic
  means to retrieve such information.
\end{itemize}

\paragraph{Intuition for our design and construction.} %
Functionally, we combine both worlds. Namely, a user can get as many credentials
as s/he wants -- although all are tied to the same user key --, and signatures
(or credential presentations) of the same credential must be unlinkable. In this
simplified first step, we restrict to selective attribute disclosure, meaning
that signatures produced by a credential reveal only a subset of the attributes
in that credential. Given such a signature, there is a trusted opener who can
de-anonymize its signer, producing back its public key, as well as a proof of
correct opening.

With respect to the model, privacy-wise this direct combination requires that
any two signatures (credential presentations) reveal nothing about the
credential used to produce them, beyond the revealed attributes. For this first
step, security-wise we stick to the conventional separation of
unforgeability-related properties in group signatures. That is, traceability
requires that, in the presence of an honest issuer, all valid signatures must
originate from a credential issued by the issuer; and, from the AC domain, we
import the requirement that no user can claim having attributes that have not
been issued to a single credential. As in group signatures, if the issuer is
corrupt, all we can hope for is that honest users do not get undeserved blame;
but, other than that, the adversary can obviously obtain credentials with
arbitrary attributes. Still, note that this is a clear gain with respect to
conventional anonymous credential systems, which offer no security against a
malicious issuer.
%
Finally, in this first step, we restrict ourselves to signatures/presentations
involving only one credential, to a system with only one issuer and opener  and,
as stated, to selective attribute disclosure.

\input{model-gsac.tex}
\input{construction-gsac.tex}
\input{security-gsac.tex}
\input{extensions-gsac.tex}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End: