### 2022-05-10 prism-backend/research/gsac/uas.tex
Improved (lengthy) abstract.

\documentclass{llncs}%[10pt]{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage{mathtools}

\title{Universal Anonymous Signatures}
\author{}

% Custom shorthands for this paper
\input{symbols.tex}
\input{symbols-gsac.tex}
\input{symbols-uas.tex}

\begin{document}
\maketitle

\begin{abstract}
  Much work has been done in the last 4 decades on privacy-preserving
  identities and authentication that tries to incorporate some sort of utility
  and accountability without losing (too much) privacy. Among the most relevant
  efforts we can find group signatures and anonymous credentials.
  %
  Anonymous credentials (AC) let users obtain attestations by trusted
  issuers on certain attributes -- thus implicitly defining the concept of
  identity as ``a set of attributes'' -- and allow credential owners to prove
  claims on these attributes. Indeed, this has proven to be very
  flexible, to the point that it has served as inspiration to design W3C's
  Verifiable Credentials, a relevant industry-led effort. However, the
  accountability offered by anonymous credentials is quite limited; frequently
  restricted to blacklisting, if present at all.
  %
  On the other hand, group signatures (GS) have traditionally focused on plain
  authentication without attributes (except very few exceptions), making instead
  more emphasis on the accountability side. That is, not only allowing full
  deanonymization, but also variants such as sequential linking, user-controlled
  anonymity, or even full unlinkability.
  %
  Not surprisingly, GS and AC share very common syntax, models and frequently
  also constructions.

  We argue that, for the real world, both functionalities are essential. Taking
  advantage of their similarities, we first propose a trivial combination, that
  we refer to with the original \GSAC acronym. Through \GSAC, we
  showcase the benefits of even a simple combination. We give a common syntax,
  security model, and generic construction for \GSAC.
  %
  Next, we acknowledge that even \GSAC has strong limitations. Namely, it is too
  strict in the rules one can apply during issuance of credentials, and signing
  and opening of the produced signatures. While generalizations along these
  lines have been explored independently in the GS and AC domains, to the best
  of our knowledge, no single scheme offers full flexibility.
  We define a syntax, security model, and give a generic
  construction for a scheme meeting these requirements. Given its high
  flexibility to cater for different needs, we name it Universal Anonymous
  Signatures (\UAS). Indeed, an \UAS scheme lets issuers decide on issuance
  policies, which easily enable advanced use cases such as delegation, or
  issuance conditioned identity-related claims. It can also be used to build
  schemes that allow signers prove arbitrary claims on their identities, as well
  as deciding on which should be the opening authority that can apply some
  commonly agreed deanonymization function over the signer's identity (which is
  now well-defined, given the incorporation of attributes from the AC domain).
  %
  The techniques and building blocks we leverage are well known. The challenge
  is, rather, on the definitional side.
\end{abstract}

\input{introduction.tex}
\input{preliminaries.tex}
\input{sec-gsac.tex}
\input{sec-uas.tex}
\input{benchmarks.tex}
\input{conclusion.tex}

\bibliographystyle{splncs04}
\bibliography{uas}

\appendix

\input{app-building-blocks.tex}
\input{app-gsac.tex}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End: