### 2022-05-11 prism-backend/research/gsac/uas.tex

\documentclass{llncs}%[10pt]{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage{mathtools}

\setcounter{tocdepth}{2}

\title{Universal Anonymous Signatures}
\author{Jesus Diaz\inst{1} \and Markulf Kohlweiss\inst{2}}

\institute{Input Output Global,\\
\email{jesus.diazvico@iohk.io},\\
\and
Input Output Global and University of Edinburgh,\\
\email{markulf.kohlweiss@ed.ac.uk}}

% Custom shorthands for this paper
\input{symbols.tex}
\input{symbols-gsac.tex}
\input{symbols-uas.tex}

\begin{document}
{\def\addcontentsline#1#2#3{}\maketitle}%\maketitle

\begin{abstract}
  Much work has been done in the last 4 decades on privacy-preserving
  identities and authentication that tries to incorporate some sort of utility
  and accountability without losing (too much) privacy. Among the most relevant
  efforts we can find group signatures (GS) and anonymous credentials (AC).
  %
  AC let users obtain attestations by trusted
  issuers on certain attributes -- thus implicitly defining the concept of
  identity as ``a set of attributes'' -- and allow credential owners to prove
  claims on these attributes. While this has proven to be very
  flexible, accountability in AC schemes is quite limited; frequently restricted
  to blacklisting, if present at all.
  %
  On the other hand, GS have traditionally focused on plain authentication
  without attributes, making instead more emphasis on the accountability side.
  That is, not only allowing full
  deanonymization, but also variants such as sequential linking, user-controlled
  anonymity, or even full unlinkability.
  %
  Not surprisingly, GS and AC share very common syntax, models and frequently
  also constructions.

  We argue that, in the real world, both functionalities are essential. Taking
  advantage of their similarities, we first propose a trivial combination,
  \GSAC, showcasing the benefits even a simple combination brings. We give a
  common syntax, security model, and generic construction for \GSAC.
  %
  Next, we generalize \GSAC into our Universal Anonymous Signatures (\UAS)
  proposal by allowing arbitrary policies at issuance, signing and opening
  time, among other improvements. While generalizations along these
  lines have been explored independently in the GS and AC domains, to the best
  of our knowledge, no single scheme offers such full flexibility.
  We define a syntax, security model, and give a generic construction for a
  scheme meeting these requirements. Furthermore, we show how our model for
  \UAS is a generalization of previous well known schemes in the GS and AC
  domains.
  %
  The techniques and building blocks we leverage are well known. The challenge
  is, rather, on the definitional side. To further evidence the universality of
  \UAS, we show how more restricted schemes can be built from vanilla \UAS.
\end{abstract}

\tableofcontents

\input{introduction.tex}
\input{preliminaries.tex}
\input{sec-gsac.tex}
\input{sec-uas.tex}
%\input{benchmarks.tex}
\input{conclusion.tex}

\bibliographystyle{splncs04}
\bibliography{uas}

\appendix

\input{app-building-blocks.tex}
\input{app-gsac.tex}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End: