### 2022-05-14 prism-backend/research/gsac/sec-uas.tex
Added fig-table sketching proof components.

\section{From \GSAC to \UAS}
\label{sec:uas}

\GSAC is a very interesting exercise that showcases the advantages of direct
combination of group signatures and anonymous credentials. However, it also
inherits a main limitation: it is too strict in regards to the utility it
offers. The most excruciating example of this fact is that in \GSAC, via \Open,
one can only get the actual identity of the signer. Even though the inclusion
of attributes makes the concept of ``identity'' meaningful, we
may not want to fully de-anonymize the signer. Instead, we may be interested in
revealing arbitrary functions of its identity and the signed message. Also, for
the sake of illustation, we restricted to the case of selective disclosure, and
to only one credential per signature. Then, the natural question is whether we
can model an extension to \GSAC that is compatible both with constructions
that support selective disclosure, and with constructions that support arbitrary
claims on the
credential attributes; and, as it is also common in the anonymous credentials
(or verifiable credentials\footnote{\url{https://www.w3.org/TR/vc-data-model/}.
  Last access, May 5th, 2022.}) domains, to allow credential holders to
combine multiple credentials into one signature/presentation. Similarly, we can
also generalize \GSAC by allowing users to leverage previously obtained
credentials in order to obtain new ones; possibly, applying arbitrary issuance
policies. In a nutshell, our generalization of \GSAC into Universal Anonymous
Signatures (\UAS), captured in \figref{fig:proof-blocks-uas}, enables:

\begin{itemize}
\item Using previously obtained credentials to request new ones, applying a
  configurable policy defined via an issuance function \fissue, which determines
  whether the new issuance should proceed or not.
\item Modulating the type of utility one wants to get directly from signatures,
  via a signature evaluation function \feval. \feval can be programmed to reveal
  nothing, selective disclosure, arbitrary claims over the credentials'
  attributes, etc.
\item Defining arbitrary accountability information that will be extractable by
  openers through an opening function \finsp. This can range from no information
  at all, to functions on subsets of the attributes, to full de-anonymization.
\end{itemize}

\begin{figure}[ht!]
  \input{figures/proof-blocks-uas.tex}
  \caption{Comparison of statements proved per operation in \GSAC and \UAS.
    $K(x)$ means proving knowledge of $x$; $E(y)$ means proving correct
    encryption of $y$; $D(z)$ means proving correct decryption of $z$. In
    \UAS, solid boxes span configurable statements that control issuance; dashed
    boxes span configurable statements that control utility; dotted boxes span
    configurable statements that control accountability. Unboxed statements are
    fixed, and are thus common to all \GSAC/\UAS schemes.}
  \label{fig:proof-blocks-uas}
\end{figure}

Furthermore, to make the concept more flexible, we no longer restrict to the
case of one issuer and one opener. Instead, \UAS supports multiple issuers (as
in generalized variants of anonymous credentials), as well as multiple openers,
each one with a predefined open function. 
A consequence of this is that the concept of ``group'', which was very present
in \GSAC, is now blurred in \UAS, and we get closer to the context of anonymous
credentials -- while still offering the accountability we inherit from \GSAC
via \Open and \Judge.
%
Although this may seem an innocent change, it brings much flexibility in
practice.

\input{model-uas.tex}
\input{construction-uas.tex}
\input{security-uas.tex}
\input{transformations-uas.tex}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End: