### 2022-06-07 prism-backend/research/gsac/security-gsac.tex

\subsection{Correctness and Security of \GSACGen}
\label{ssec:security-gsac}

For the sake of brevity, here we just enuntiate the main theorems related to
correctness and security of \GSAC, and defer the full proofs to
\appref{sapp:gsac-proofs}.

\begin{theorem}[Correctness of \GSACGen]
  \label{thm:correctness-gsac}
  If the underlying \SBCM, encryption, and commitment schemes are correct, and
  the NIZK systems for $NIZKRel_{\Issue}$, $NIZKRel_{\Sign}$, and $NIZKRel_{\Open}$
  are also correct, our generic construction \GSACGen
  satisfies correctness as defined in \defref{def:correctness-gsac}.
\end{theorem}

\begin{theorem}[Issuance anonymity of \GSACGen]
  \label{thm:iss-anonymity-gsac}
  \todo{XXX}
\end{theorem}

\begin{theorem}[Signature anonymity of \GSACGen]
  \label{thm:sig-anonymity-gsac}
  If the underlying $\NIZK^{\NIZKRel_{\Sign}}$ scheme is zero-knowledge and
  simulation-extractable, and the encryption system is IND-CCA secure,
  our \GSACGen construction satisfies anonymity as defined in
  \defref{def:sig-anonymity-gsac}.
\end{theorem}

\begin{theorem}[Traceability of \GSACGen]
  \label{thm:trace-gsac}
  If the underlying $\NIZK^{\NIZKRel_{\Sign}}$ scheme is sound, the
  $\NIZK^{\NIZKRel_{\Open}}$ is  complete, the public-key encryption algorithm
  is correct, and the signature scheme for blocks of committed messages is
  one-more unforgeable, then our \GSACGen construction satisfies traceability as
  defined in \defref{def:trace-gsac}, except with negligible
  probability.
\end{theorem}

\begin{theorem}[Non-frameability of \GSACGen]
  \label{thm:frame-gsac}
  If the NIZK schemes for $\NIZKRel_{\Issue}$ is zero-knowledge, the scheme for
  $\NIZKRel_{\Sign}$ is zero-knowledge and sound, the scheme for $\NIZKRel_{\Open}$ is
  sound, and the commitment scheme \C is hiding, then our generic \GSACGen
  construction satisfies non-frameability as defined in \defref{def:frame-gsac},
  except with negligible probability.
\end{theorem}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "gsac"
%%% End: