### 2022-06-18 prism-backend/research/gsac/conclusion-paper.tex
Beginning refurbishment of the paper.

\section{Conclusion}
\label{sec:conclusion}

[1 page max]

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas-paper"
%%% End: