### 2022-06-18 prism-backend/research/gsac/uas-paper.tex

\documentclass{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

%% Set/unset to activate/deactivate author comments
\def\authnotes{1}

%%% Allows line-breaks in math formulae after commas
\AtBeginDocument{%
  \mathchardef\mathcomma\mathcode`\,
  \mathcode`\,="8000 
}
{\catcode`,=\active
  \gdef,{\mathcomma\discretionary{}{}{}}
}

%%%

\setcounter{tocdepth}{2}

\title{Universal Anonymous Signatures}
\author{Jesus Diaz\inst{1} \and Markulf Kohlweiss\inst{2}}

\institute{Input Output Global,\\
\email{jesus.diazvico@iohk.io},\\
\and
Input Output Global and University of Edinburgh,\\
\email{markulf.kohlweiss@ed.ac.uk}}

% Custom shorthands for this paper
\input{symbols.tex}
\input{symbols-gsac.tex}
\input{symbols-uas.tex}

\begin{document}
{\def\addcontentsline#1#2#3{}\maketitle}%\maketitle

\begin{abstract}
  \todo{XXX}
\end{abstract}

\input{introduction-paper.tex}
\input{preliminaries-paper.tex}
\input{formal-uas.tex}
\input{relationships.tex}
\input{conclusion-paper.tex}

\bibliographystyle{splncs04}
\bibliography{uas-paper}

% \appendix

% \input{app-building-blocks.tex}
% \input{app-gsac.tex}
% \input{app-model-related.tex}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End: