### 2022-06-27 prism-backend/research/gsac/conclusion-paper.tex
Added a paragraph with real-world sample use cases. Fixed some todos.

\section{Conclusion}
\label{sec:conclusion}

\todo{Try to restrict to 0.5 pages max}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas-paper"
%%% End: