### 2022-08-11 prism-backend/research/gsac/app-related-proofs.tex
Reviewed proofs of related schemes up to AC. Added ref to MPS eprint.

\section{Proofs for Relationships with Other Schemes}
\label{app:related-proofs}

\subsection{Security of \CUASGS}

We prove that our \CUASGS construction is a secure group signature scheme,
according to \cite{bsz05}, if the underlying \CUASGen construction is secure.

\begin{theorem}[Anonymity of \CUASGS]
  If the base \CUASGen construction has signature anonymity according to
  \defref{def:sign-anonymity-uas}, then \CUASGS is an anonymous group
  signature scheme.
\end{theorem}

\begin{proof}
  Assume an adversary \adv~ against anonymity in \CUASGS. This directly leads
  to an adverary \advB against anonymity of the $(\fissue^1,\feval^0,
  \finsp^{\upk})-\CUASGen$ instance.

  Concretely, observe that calls to \adv's $Ch,Open,SndToU,WReg,USK,CrptU$
  oracles can be directly simulated by calls to \advB's oracles $\CHALb,
  \OPEN,\HUGEN,\WREG,\CUGEN,\UCORR$ oracles. Thus, if \adv~guesses its bit $b$
  with non-negligible probability, this directly leads to a non-negligible
  probability of guessing \advB's challenge bit.
  \qed
\end{proof}

\begin{theorem}[Traceability of \CUASGS]
  If the base \CUASGen construction is unforgeable according to
  \defref{def:forge-uas}, then \CUASGS is a traceable group signature
  scheme.
\end{theorem}

\begin{proof}
  Assume an adversary \adv~against traceability in \CUASGS. We build an
  adversary \advB against unforgeability of the $(\fissue^1,\feval^0,
  \finsp^{\upk})-\CUASGen$ restriction.

  \advB's simulation of \adv's environment is as in the proof for anonymity
  of \CUASGS. \advB simply outputs whatever \adv~outputs in line 2 of its
  traceability game. Note that, whenever \CUASGS accepts a signature, so does
  \CUASGen. Concretely,
  assume that \adv~wins the traceability game in \figref{fig:model-gs} with
  $\upk \notin \HU \cup \CU$ at line 5. This implies that, in line 14 of the
  \UAS unforgeability game (\figref{fig:exp-uas-unfor}) $\OWNR[\lbrace \cid_i
  \rbrace_{i\in[n]}]$ returns an \uid which is not in either \HU or \CU.
  Similarly, if \adv~wins the traceability game with condition $\Judge(\dots)=0$
  at line 5, then \advB wins its unforgeability game with the corresponding
  check at line 6 in \figref{fig:exp-uas-unfor}.
  %
  \qed
\end{proof}

\begin{theorem}[Non-frameability of \CUASGS]
  If the base \CUASGen construction is non-frameable according to
  \defref{def:frame-uas}, then \CUASGS is a non-frameable group signature
  scheme.
\end{theorem}

\begin{proof}
  Assume an adversary \adv~against non-frameability in \CUASGS. We build an
  adversary \advB against non-frameability of the $(\fissue^1,\feval^0,
  \finsp^{\upk})-\CUASGen$ instance.

  \advB's simulation of \adv's~environment is as in the proof for anonymity
  of \CUASGS. \advB simply outputs whatever \adv~outputs at line 2 of its
  game. Observe that the winning condition for \adv~directly implies a winning
  condition for \advB. Concretely, if \adv~wins its game, \advB wins its game
  in line 11 of \ExpNonframe.
  \qed
\end{proof}

\subsection{Security of \CUASRing}

We prove that our \CUASRing construction is anonymous and unforgeable, if
the underlying $\CUASGen^{hide-iss}$ construction is secure.

\begin{theorem}[Anonymity of \CUASRing]
  If the base $\CUASGen^{hide-iss}$ construction is anonymous according to
  \defref{def:sign-anonymity-uas}, then \CUASRing is an anonymous ring signature
  scheme.
\end{theorem}

\begin{proof}
  Assume $\adv$ is an adversary against ring signature anonymity for \CUASRing.
  Then, we build \advB against anonymity of \CUASGen. Concretely, \advB first
  generates $n$ key pairs for \adv simulating $KeyGen$ in \adv's game. Since
  $KeyGen$ simply uses algorithms of $\CUASGen^{hide-iss}$, \advB can do this
  directly via its own \HUGEN, \ISET, and \OGEN oracles, specifying
  $\fissue^{\sring}$ as issuance function, and $\finsp^0$ as opening function.
  Then, \advB invokes $\adv$ passing the $n$ produced public keys, and
  answers $\adv$'s $SIGN(i,R,\msg)$ queries as follows. If user $i$ has not yet
  produced a credential for ring $\pk_i \in R$, \advB makes a call to its
  \OBTAIN oracle, using $R$ as attribute set. If such a credential was already
  produced, \advB simply reuses it. Then, \advB uses the corresponding user and
  credential identifiers as parameters to a query to its own \SIGN oracle
  (specifying $\feval^{\attrs}$ as signature evaluation function). Eventually,
  $\adv$ outputs two user indexes, a ring $R$, and message \msg. If the checks
  at lines 3-4 of \figref{fig:model-rs} pass, \advB generates the challenge
  signature by calling its own \CHALb oracle, and passes the result to $\adv$.
  Finally, \advB outputs whatever $\adv$ outputs.
  %
  The simulation of \adv's anonymity game is perfect, and \advB wins its game
  whenever \adv does.
  \qed
\end{proof}

\begin{theorem}[Unforgeability of \CUASRing]
  If the base $\CUASGen^{hide-iss}$ construction is non-frameable according to
  \defref{def:frame-uas}, then \CUASRing is an unforgeable ring signature
  scheme.
\end{theorem}

\begin{proof}
  Assume $\adv$ against unforgeability of \CUASRing wins the game. We build
  \advB winning the non-frameability game of \CUASGen with the same probability.
  %
  \advB prepares the environment as the adversary \advB against anonymity,
  and answers \adv's queries to the $SIGN$ oracle also as in the anonymity
  proof. To answer \adv's queries to its $CORR$ oracle, \advB leverages the
  \ICORR oracle in the non-frameability game for \UAS. Finally, \advB outputs
  whatever $\adv$ outputs.

  Note that, if $\adv$ wins its game, it means that the public keys of the
  issuers in $R$ all belong to uncorrupted users. Since the signature is
  accepted by $Verify$ at line 3 of the ring signature unforgeability game,
  it is also accepted by \Verify in the \UAS non-frameability game. But,
  moreover, given the construction of $Verify$ in \CUASRing, there is some
  honest \usk within $R$ that produced the signature, so $\OWNR[\scid] \in \HU$
  in \advB's non-frameability game. Also, since $\adv$ never queried its $SIGN$
  oracle, then by construction there cannot exist a matching entry in the
  $\SIG[\uid]$ table. Thus, \advB wins the non-frameability game whenever
  $\adv$ wins its unforgeability game. Since \CUASGen is non-frameable, then
  this probability must be negligible.
  %
  \qed
\end{proof}

\subsection{Security of \CUASAC}

We prove that \CUASAC is an anonymous and unforgeable anonymous credential
scheme, according to \cite{fhs19}, if the underlying \CUASGen construction
is anonymous and unforgeable. Note that AC schemes modelled as in \cite{fhs19}
do not have issuance anonymity, as the user public key is passed to the issuer.

\begin{theorem}[Anonymity of \CUASAC]
  \label{thm:anon-cuasac}
  If the base \CUASGen construction has signature anonymity according to
  \defref{def:sign-anonymity-uas}, then \CUASAC is an anonymous AC scheme
  according to \cite{fhs19}.
\end{theorem}

\begin{proof}
  Given \adv~against anonymity of \CUASAC as defined in \cite{fhs19}, we build
  an adversary \advB against anonymity of \CUASGen as defined in
  \defref{def:sign-anonymity-uas}.

  To simulate oracle calls by \adv, \advB simply redirects to its corresponding
  oracle, i.e.: for calls to $HU,CU,Obtain,Show,LoR$, \advB redirects to
  \HUGEN,\CUGEN,\OBTAIN,\SIGN,\CHALb, respectively. Note that ``translating''
  the inputs from the former into inputs for the latter is trivial.
  Obviously, if \adv~wins in its anonymity game, then so does \advB with the
  same probability.
  % 
  \qed
\end{proof}

\begin{theorem}[Unforgeability of \CUASAC]
  \label{thm:forge-cuasac}
  If the base \CUASGen construction is unforgeable according to
  \defref{def:forge-uas}, then \CUASAC is an unforgeable AC scheme.
\end{theorem}

\begin{proof}
  Given \adv~against unforgeability of \CUASAC as defined in \cite{fhs19}, we
  build an adversary \advB against unforgeability of \CUASGen as defined in
  \defref{def:forge-uas}.

  To simulate oracle calls by \adv, \advB acts as in the anonymity proof
  ($ObtIss$ and $Issue$ for \adv~correspond to \OBTISS and \ISSUE for \advB).
  Assume that \adv~wins its game. Note that, since $b=1$ in line 4 of the
  unforgeability game of \figref{fig:model-ac}, then by construction, \Verify
  also returns $1$ in the unforgeability game of \UAS. In addition, if
  there is no corrupt user with a credential containing $D$, then this leads to
  \advB winning at line 12 in \figref{fig:exp-uas-unfor}, as no credential
  owned by a corrupt user that can be extracted can make the $\feval^{\dattrs}$
  function output $\yeval^0=D$. Thus, \advB wins with the same probability as
  \adv~does.  
  %
  \qed
\end{proof}

\subsection{Security of \CUASRAC}

We prove that \CUASRAC is an anonymous and unforgeable anonymous credential
scheme, according to \cite{fhs19}, if the underlying \CUASGen construction
is anonymous and has unforgeable signing. Note that issuance unforgeability
of \CUASGen is not needed, as the model in \cite{fhs19} does not allow
leveraging other credentials not any fixed custom issuance policy. Thus,
signature unforgeability from \UAS is enough. Note that AC schemes modelled
as in \cite{fhs19} do not have issuance anonymity, as the user public key
is passed to the issuer.

\begin{theorem}[Anonymity of \CUASRAC]
  \label{thm:anon-cuasrac}
  If the base \CUASGen construction has signature anonymity according to
  \defref{def:sign-anonymity-uas}, and $F$ is one-way and collision-resistant,
  then \CUASAC is an anonymous AC scheme according to \cite{fhs19}.
\end{theorem}

\begin{proof}
  Given \adv~ against anonymity of \CUASRAC as defined in
  \figref{fig:model-rac}, we build an adversary \advB against anonymity of
  \CUASGen as defined in \defref{def:sign-anonymity-uas}.

  First observe that, given a revocation handle $F(V,id)$ extracted from any
  credential showing, it is not possible to extract $id$, as $F$ is one-way
  and collision-resistant.

  To simulate oracle calls by \adv, \advB simply redirects to its corresponding
  oracle, i.e.: for calls to $HU,CU,Obtain,Show,LoR,ReqCredHandle$, \advB
  redirects to \HUGEN, \CUGEN, \OBTAIN, \SIGN, \CHALb, of \OPEN respectively.
  Note that ``translating'' the inputs from the former into inputs for the
  latter is trivial. Also, observe that $ReqCredHandle$ essentially refuses to
  open transcripts from challenge users (signatures, in the non-interactive
  equivalent), as $F(V,\attrs_1(\ccid_0) \neq F(V,\attrs_1(\ccid_1)$ with
  overwhelming probability (see the $\tyinsp \neq \yinsp$) check in
  \figref{fig:oracles2}. Obviously, if \adv~ wins in its anonymity game, then
  so does \advB with the same probability.
  % 
  \qed
\end{proof}

\begin{theorem}[Unforgeability of \CUASRAC]
  If the base \CUASGen construction has sign unforgeability according to
  \defref{def:sign-forge-uas}, and $F$ is one-way and collision-resistant, then
  \CUASRAC is an unforgeable RAC scheme.
\end{theorem}

\begin{proof}
  Given \adv~ against unforgeability of \CUASRAC as defined in
  \figref{fig:model-rac}, we build an adversary \advB against sign
  unforgeability of \CUASGen as defined in \defref{def:sign-forge-uas}.

  To simulate oracle calls by \adv, \advB acts as in the anonymity proof. Assume
  that \adv~ wins its game via the winning condition of line 6 in the RAC
  unforgeability game. The same reasoning as in \CUASAC applies here. Next, note
  that the probability of \adv~ winning its game via the winning condition at
  line 9 is negligible, due to soundness of the NIZK scheme used for signing,
  and the one-wayness and collision-resistance of $F$ in
  $\feval^{\dattrs,F,V,L}$ and $\finsp^{F}$.
  % 
  \qed
\end{proof}

\subsection{Security of \CUASMPS}

As stated, we restrict to privacy-1 of MPS. Unforgeability-1 in MPS is
equivalent to sign unforgeability in \UAS. Unforgeability-2 in MPS is
equivalent to non-frameability in \UAS. Finally note that, in UAS, we do not
explicitly model extractability, but rather make it an inherent requirement in
the unforgeability and non-frameability definitions. Thus, our \CUASMPS
construction also ensures MPS-extractability.

\begin{theorem}[Privacy-1 of \CUASMPS]
  If the base \CUASGen construction has signature anonymity according to
  \defref{def:sign-anonymity-uas}, then \CUASMPS satisfies privacy-1
  according to \cite{ngsy22}.
\end{theorem}

\begin{proof}
  The proof is direct, by observing that the models are equivalent.

  First, note that, in MPS the adversary is required to output a function $F$,
  challenge message, and a pair of user secret keys and witnesses (in \UAS, the
  witnesses are the credentials). This is precisely what \CHALb requires as
  input in our \UAS modelling.

  Next, in MPS, validity of the output of $F$ is checked. In MPS, this
  translates to checking that $F$ does not ouptut $0$. In UAS, this is checked
  in the \SIGN and \CHALb oracles, which abort if the output produced by \feval
  does not belong in \rngfeval -- a well defined set of acceptable outputs by
  the chosen \feval.

  The MPS adversary is then allowed to make more oracle queries, and outputs a
  response bit. Just as in \UAS. Thus, if \CUASMPS satisfies signature
  anonymity, it is also an MPS scheme satisfying privacy-1.  
  % 
  \qed
\end{proof}

\begin{theorem}[Extractability of \CUASMPS]
  If the base \CUASGen construction has signature unforgeability according to
  \defref{def:sign-forge-uas}, then it is an extractable MPS scheme.
\end{theorem}

\begin{proof}
  The proof is direct, as the conditions checked in the extractability
  experiment in MPS are also checked in the signature unforgeability experiment
  for \UAS.

  Concretely, \Verify must accept the signature, and the ouptut of \feval and
  \finsp, computed with the values extracted by \ExtractSign, must respectively
  match those output by the adversary (for \feval), and computed honestly by the
  signature unforgeability (for \finsp).  
  \qed
\end{proof}

\begin{theorem}[Unforgeability-1 of \CUASMPS]
  If the base \CUASGen construction has signature unforgeability according to
  \defref{def:sign-forge-uas}, then \CUASMPS satisfies unforgeability-1 as
  defined in \cite{ngsy22}.
\end{theorem}

\begin{proof}
  For an MPS scheme to satisfy type-1 unforgeability, it has to be extractable.
  This is directly satisfied by signature unforgeability in \UAS.
  %
  Next, the adversary wins in the unforgeability-1 MPS experiment if it manages
  to produce a signature that has not been produced via a call to their \SIGN
  oracle, nor it can be associated to an corrupt user. Indeed, if that happens,
  then such adversary can be used to build an adversary winning the signature
  unforgeability game against \CUASMPS, through winning condition at line 12 --
  as \IdentifySig does not return 1 for any known \uid. Thus, if \CUASMPS has
  signature unforgeability, it also satisfies unforgeability-1 of MPS schemes.
  %
  \qed
\end{proof}

\begin{theorem}[Unforgeability-2 of \CUASMPS]
  If the base \CUASGen construction has non-frameability according to
  \defref{def:frame-uas}, then \CUASMPS satisfies unforgeability-2 as
  defined in \cite{ngsy22}.
\end{theorem}

\begin{proof}
  For an MPS scheme to have type-2 unforgeability, it has to be extractable.
  This is directly satisfied by non-frameability in UAS (a similar reasoning
  as for signature unforgeability and extractability applies here).
  %
  Next, in MPS, the adversary wins if the produced signature has not been
  output by the \SIGN oracle, but Extract outputs the identifier of an honest
  user. This is the same check done in lines 9 and 10 of the non-frameability
  game in \UAS. Therefore, if \CUASMPS satisfies non-frameability as defined in
  \defref{def:frame-uas}, it also satisfies type-2 unforgeability in MPS.
  \qed
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas-paper"
%%% End: