### 2022-08-17 prism-backend/research/gsac/uas-paper.tex

\documentclass{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
%\usepackage{mathusertools}
\usepackage{multirow}
\usepackage{footnote}
\usepackage{framed}
\usetikzlibrary{decorations.pathreplacing,calligraphy}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

%% Set/unset to activate/deactivate author comments
\def\authnotes{0}

%%% Allows line-breaks in math formulae after commas
\AtBeginDocument{%
  \mathchardef\mathcomma\mathcode`\,
  \mathcode`\,="8000 
}
{\catcode`,=\active
  \gdef,{\mathcomma\discretionary{}{}{}}
}

%%%

\setcounter{tocdepth}{2}

\title{Universal Anonymous Signatures}
\author{Jesus Diaz\inst{1} \and Markulf Kohlweiss\inst{2}}

\institute{Input Output Global,\\
\email{jesus.diazvico@iohk.io},\\
\and
Input Output Global and University of Edinburgh,\\
\email{markulf.kohlweiss@ed.ac.uk}}

% Custom shorthands for this paper
\input{symbols.tex}
\input{symbols-gsac.tex}
\input{symbols-uas.tex}

\begin{document}
{\def\addcontentsline#1#2#3{}\maketitle}%\maketitle
\ifnum\authnotes=1
Version of \today
\fi

\begin{abstract}
  For decades, anonymous credentials, group signatures, and closely related
  systems in the domain of privacy-preserving authentication, have relied on
  common cryptographic building blocks. However, each has its own approach
  modelling-wise, despite obvious commonalities.
  %
  Even within systems of the same type, multiple ``sub-models'' coexist that
  cater for slightly different requirements.

  In this work, we make a significant generalization effort in order to come
  up with a flexible game-based model that subsumes (a reasonable subset of)
  previous models in this domain. The key to this flexibility lies in functional
  placeholders that, with careful constraints, still allow to capture the
  usual anonymity, unforgeability and non-frameability properties. We refer
  to this unified model as \emph{Universal Anonymous Signatures} (\UAS).

  To showcase the generality of our model, we first give a generic construction
  of \UAS, along with minor variations of this construction that also meet the
  \UAS model. Then, we prove that well-known related schemes can be built as a
  special case (restriction) of our generic construction, just by giving
  concrete definitions of the previously mentioned functional placeholders.
  In particular, we prove our \UAS model spans (vanila) group signatures, group
  signatures with message-dependent opening, multimodal private signatures,
  attribute-based anonymous credentials, revocable anonymous credentials,
  delegatable anonymous credentials, and ring signatures. This also proves
  interesting (and, in some sense, folklore) facts, such as the fact that ring
  signatures can be seen as a multi-issuer group signature scheme, in which
  every user is also an issuer. Also, it paves the way to come up with
  ``hybrid'' constructions where one (e.g.) combines features of anonymous
  credentials (proving predicates at authentication time), with features of
  group signatures (conditional deanonymization based on attributes of the
  signer).

  Besides the main theoretical contribution of providing a unified model for
  many related schemes, we do believe that a (possibly more relevant) practical
  contribution consists on the fact that system designers and cryptographic
  engineers may adotp some generic construction of \UAS, and then simply focus
  on concrete definitions of a few functional placeholders, to tweak the
  different utility-vs-privacy tradeoff they want to achieve -- without the need
  to prove everything from scratch, nor needing to switch to a completely
  different construction if their needs change.
\end{abstract}

\input{introduction-paper.tex}
\input{preliminaries-paper.tex}
\input{formal-uas.tex}
\input{relationships.tex}
\input{conclusion-paper.tex}

\bibliographystyle{splncs04}
\bibliography{uas-paper}

\appendix

\input{app-building-blocks.tex}
\input{app-uas-proofs.tex}
%\input{app-interactive.tex}
\input{app-related-proofs.tex}
\input{app-related-extra.tex}
% \input{app-gsac.tex}
% \input{app-model-related.tex}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas-paper"
%%% End: