### 2022-08-17 prism-backend/research/gsac/uas-paper.tex
Refining

\documentclass{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
%\usepackage{mathusertools}
\usepackage{multirow}
\usepackage{footnote}
\usepackage{framed}
\usetikzlibrary{decorations.pathreplacing,calligraphy}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

%% Set/unset to activate/deactivate author comments
\def\authnotes{0}

%%% Allows line-breaks in math formulae after commas
\AtBeginDocument{%
  \mathchardef\mathcomma\mathcode`\,
  \mathcode`\,="8000 
}
{\catcode`,=\active
  \gdef,{\mathcomma\discretionary{}{}{}}
}

%%%

\setcounter{tocdepth}{2}

\title{Universal Anonymous Signatures}
\author{Jesus Diaz\inst{1} \and Markulf Kohlweiss\inst{2}}

\institute{Input Output Global,\\
\email{jesus.diazvico@iohk.io},\\
\and
Input Output Global and University of Edinburgh,\\
\email{markulf.kohlweiss@ed.ac.uk}}

% Custom shorthands for this paper
\input{symbols.tex}
\input{symbols-gsac.tex}
\input{symbols-uas.tex}

\begin{document}
{\def\addcontentsline#1#2#3{}\maketitle}%\maketitle
\ifnum\authnotes=1
Version of \today
\fi

\begin{abstract}
  For decades, anonymous credentials, group signatures, and closely related
  systems in the domain of privacy-preserving authentication, have relied on
  common cryptographic building blocks. However, each has its own approach
  modelling-wise, despite obvious commonalities.
  %
  Even within systems of the same type, multiple ``sub-models'' and
  constructions coexist that cater for slightly different requirements.

  Previous works, like Camenisch et al in SAC'15, or Nguyen et al in CRYPTO'22
  generalize aspects of those systems, but still within their own domain --
  anonymous credentials and group signatures, respectively.

  In this work, we make a significant generalization effort in order to come
  up with a flexible game-based model that subsumes
  previous ones. The key to this flexibility lies in functional placeholders
  that, with careful constraints, still allow to capture generalized notions of
  the usual anonymity, unforgeability and non-frameability properties. We refer
  to this unified model as \emph{Universal Anonymous Signatures} (\UAS).

  To showcase the generality of \UAS, we give a main construction, plus two
  alternative ones that slightly modify it, following our \UAS model.
  We prove that, from these constructions, we can build multiple variants of
  group signatures, anonymous credentials, and ring signatures, just by giving
  concrete definitions of the mentioned functional placeholders.

  Our main theoretical contribution is unifying previously separate
  (albeit close) systems under the same generalized model. Indeed, such unified
  model allows to, very easily, build ``hybrid'' systems that combine features
  of previously separately modelled constructions: e.g., anonymous credentials
  with conditional deanonymization based on their attributes.

  More practically, this has the consequence that cryptographic
  engineers and system designers can tweak the desired utility-vs-privacy
  tradeoff that best suits their needs, just by defining a few functions --
  but without needing to prove security from scratch or implement
  a whole system.
\end{abstract}

\input{introduction-paper.tex}
\input{preliminaries-paper.tex}
\input{formal-uas.tex}
\input{relationships.tex}
\input{conclusion-paper.tex}

\bibliographystyle{splncs04}
\bibliography{uas-paper}

\appendix

\input{app-building-blocks.tex}
\input{app-uas-proofs.tex}
%\input{app-interactive.tex}
\input{app-related-proofs.tex}
\input{app-related-extra.tex}
% \input{app-gsac.tex}
% \input{app-model-related.tex}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas-paper"
%%% End: