### 2022-08-31 prism-backend/research/gsac/status.tex

\section{Where Are We?}
\label{sec:status}

Currently, we are completing our first iteration of foundational research. We
have a formal security model, generic construction, and security proofs for our
construction. This means that we have a solid common ground from which we can
start building concrete instantiations of UAS (such as those in
\secref{sec:use-cases}), and evaluate their performance in the real world.

Moreover, we are also beginning to work on integrating UAS within Atala Prism%
\footnote{\url{https://atalaprism.io}. Last access, August 30th, 2022.}, the
Self-Sovereign Identity (SSI) product from IOG\footnote{\url{https://iog.io}.
  Last access, August 30th, 2022.}, built atop of Cardano%
\footnote{\url{https://cardano.org}. Last access, August 30th, 2022.}, and
compliant with the most relevant specifications in the SSI domain.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas-onepager"
%%% End: