### 2022-10-03 prism-backend/research/gsac/uas-paper.tex

\documentclass{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
%\usepackage{mathusertools}
\usepackage{multirow}
\usepackage{dashbox}

% Box types
\newcommand\dboxed[1]{\dbox{\ensuremath{#1}}}
\makeatletter
\newcommand{\dotboxed}[1]{%
\tikz[baseline={([yshift=-1ex]current bounding box.center)}] \node [rectangle, minimum width=1ex,dotted,draw,line width=0.6pt] {\normalcolor\m@th$\displaystyle#1$};}
\makeatother
\usepackage{footnote}
\usepackage{framed}
\usetikzlibrary{decorations.pathreplacing,calligraphy}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

%% Set/unset to activate/deactivate author comments
\def\authnotes{0}

%%% Allows line-breaks in math formulae after commas
\AtBeginDocument{%
  \mathchardef\mathcomma\mathcode`\,
  \mathcode`\,="8000 
}
{\catcode`,=\active
  \gdef,{\mathcomma\discretionary{}{}{}}
}

%%%

\setcounter{tocdepth}{2}

\title{Universal Anonymous Signatures}
\author{Jesus Diaz\inst{1} \and Markulf Kohlweiss\inst{2}}

\institute{Input Output Global,\\
\email{jesus.diazvico@iohk.io},\\
\and
Input Output Global and University of Edinburgh,\\
\email{markulf.kohlweiss@ed.ac.uk}}

% Custom shorthands for this paper
\input{symbols.tex}
\input{symbols-gsac.tex}
\input{symbols-uas.tex}

\begin{document}
{\def\addcontentsline#1#2#3{}\maketitle}%\maketitle
\ifnum\authnotes=1
Version of \today
\fi

\begin{abstract}
  For decades, anonymous credentials (AC), group signatures (GS), and
  related schemes in the field of privacy-preserving authentication, have
  relied on common cryptographic tools. Yet, each has its own
  approach modelling-wise. Even within systems of the same type, multiple
  ``sub-models'' coexist, catering for slightly different needs.
  %
  Works like Camenisch et al (SAC'15) or Nguyen et al (CRYPTO'22)
  generalise some aspects, but still within their own domain -- AC and GS,
  respectively.

  In this work, we make a further effort in order to come up with a flexible
  game-based model that generalises and unifies GS and AC schemes. The key to
  this flexibility lies in functional placeholders that, with careful
  constraints, allow to capture generalised notions of the usual anonymity,
  unforgeability and non-frameability properties. We refer to this unified model
  as \emph{Universal Anonymous Signatures} (\UAS).

  To showcase the generality of \UAS, we give a construction following our model
  and prove that, from it, we can easily build multiple variants of group
  signatures, anonymous credentials, and even ring signatures, just by giving
  concrete definitions of the mentioned functional placeholders.

  Our main theoretical contribution is thus unifying previously separate (albeit
  close) systems under the same generalised model. Such unified model also
  allows to, very easily, build ``hybrid'' systems combining features of its
  ancestors: e.g., anonymous credentials with conditional deanonymization based
  on their attributes.

  In practice, this means that cryptographic engineers and system designers
  can reach the desired utility-vs-privacy tradeoff that best suits their needs,
  just by defining a few functions -- but without needing to model and prove
  security from scratch or implement a whole system.
\end{abstract}

\input{introduction-paper.tex}
\input{preliminaries-paper.tex}
\input{formal-uas.tex}
\input{relationships.tex}
\input{conclusion-paper.tex}

\bibliographystyle{splncs04}
\bibliography{uas-paper}

\appendix

\input{app-building-blocks.tex}
\input{app-uas-proofs.tex}
%\input{app-interactive.tex}
\input{app-related-proofs.tex}
\input{app-related-extra.tex}
% \input{app-gsac.tex}
% \input{app-model-related.tex}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas-paper"
%%% End: