### 2022-11-03 prism-backend/research/prism-protocol/vc_atala.tex

\section{VCs in Atala PRISM}
\label{sec:vc-atala}

In Atala PRISM, VCs are associated to a DID, with no limit on the number of VCs
per DID. VCs (or just, credentials) can be issued, revoked, and presented --
where a presentation involves both proving and verifying steps.

\subsubsection{VC structure.}
A VC contains at least the DID and issuing key index of the issuer, the DID and
authentication key index of the holder, and a dictionary with further identity
claims (name, nationality, etc.) This data is then signed by the issuing key.
A credential presentation requires the holder to prove control over the
specified authentication key. Issuers are trusted to verify the identity
claims contained in credentials they issue.

\subsubsection{Issuing VCs.}
Issuing a VC involves creating and signing the corresponding data structure, and
publishing it in the blockchain. In Atala PRISM, the publication can be done in
batches, so in the following we treat all publication events as containing a
batch of VCs (of size at least $1$).

The protocol for creating a single VC is as defined in \figref{fig:issuevc}.
Therein, we only describe how the prospective VC holder proves to the issuer
ownership of the DID to be included in the VC. Indeed, this is the only
claim whose correctness can be proved cryptographically; whereas for the other
claims we need to trust the issuer. \figref{fig:issuevc} further captures how
the issuer batches multiple VCs and publishes them in the blockchain.

\begin{figure}
  \scalebox{0.8}{
    \procedure{\VCCreate}{%
      \text{$\mathbf{Holder}(\did_H,\did_I,\ask)$} \> \>
      \text{$\mathbf{Issuer}(\did_I,\isk)$} \> \> \textbf{VDR} \\
      \> \sendmessageright{top=\text{$\did_H$,attrs}} \> \> \> \\
      \> \> \> \sendmessageright{top=$\DID.Read(\did_H)$} \> \\
      \> \> \> \sendmessageleft{top=\apk} \> \\
      \> \> r \getr \ZZ \> \> \\
      \> \sendmessageleft{top=$r$} \> \> \> \\
      \sig \gets \Sig.Sign(\ask,r) \> \> \> \> \\
      \> \sendmessageright{top=\sig} \> \> \> \\
      \> \> \pcif !\Sig.Verify(\apk,\sig,r): \> \> \\
      \> \> \pcind \pcreturn \bot \> \> \\
      \> \> c \gets (\ipk,\apk,attrs) \> \> \\ 
      \> \> vc \gets \Sig.Sign(\isk,c) \> \> \\
      \> \sendmessageleft{top=\text{$(c,vc)$}} \> \> \> \\  
      \pcif !\Sig.Verify(\ipk,vc,c): \> \> \> \> \\
      \pcind \pcreturn \bot \> \> \> \> \\
    }
  }
  \label{fig:issuevc}
  \caption{Algorithms for issuing (creating and publishing) a VC.}
\end{figure}

\subsubsection{Revoking a VC.}

\subsubsection{VC Presentations.}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "prism-protocol"
%%% End: