### 2022-12-17 prism-backend/research/prism-protocol/gpkivdu.tex
Better defined PKI functionality.

\section{A PKI with Verifiable Delayed Updates}
\label{sec:pkivdu}

In this section, we present the ideal functionality for a PKI that supports
updates which are:

\begin{itemize}
\item Verifiable: Any party with read access to the PKI storage can check
  that the last state of any stored public key has been produced after
  applying a sequence of authenticated changes to this public key.
\item Delayed: Any party with write access to the PKI storage is guaranteed
  that its requested changes will get applied after at most some predefined
  delayed.
\end{itemize}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "prism-protocol"
%%% End: