### 2022-12-17 prism-backend/research/prism-protocol/pkivdu.tex

\section{\IdealGPKIDID: A PKI with Verifiable Delayed Updates}
\label{sec:pkivdu}

In this section, we present the ideal functionality for a PKI that supports
updates which (informally) are:

\begin{itemize}
\item Verifiable: It can be checked that the sequence of changes that lead to
  the current state of a (set of) public key(s) have been applied by their
  legitimate owner.
\item Delayed: Changes to existing (sets of) public keys, or requests to add
  new ones, get applied after at most some predefined delay.
\end{itemize}

\paragraph{The ideal functionality.} %
We take the generic approach to see a PKI as a system that stores dictionaries
(i.e., lists of labels with corresponding types and values), where each
dictionary is associated to a single identifier. The functionality is then
defined in terms of the (CRUD) operations that can be run to manage such
objects. To define the ideal functionality as abstractly as possible, we only
make use of one external (global) subroutine: a clock \IdealGclock.
Indeed, given that the objects stored in the PKI can be modified, and such
objects are in fact public keys used to authenticate entities at concrete points
in time, keeping track of when a mocification takes place, is crucial -- e.g.,
to know if a public key is/was valid at a certain point in time.
%
Otherwise, the functionality simply keeps track of the requested (CRUD)
operations, makes sure that only the entitled party can perform them, and that
the maximum delay is not violated. Note that this delay can be set to any
arbitrarily small value, which means that a delayed PKI is simply a
generalisation of a (non-delayed) PKI.
%
Finally, we parameterize the ideal functionality with a policy \P, that we use
to account for any arbitrary requirements that concrete instantiations may
impose on the contents of the objects managed by the PKI. For instance, some
concrete PKI may refuse to store RSA public keys of length less than $4096$
bits; or, in the DID setting, a DID method may require that all DID Documents
contain an active key of some concrete type.

We now describe the interfaces exposed by \IdealGPKIDID, following the previous
guidelines. They are formally defined in \figref{fig:fpkidid}.

\begin{description}
\item[Create.] Upon receiving the instruction \uccmd{Create}, the functionality
  adds a new set of values $\val_i$, each of type $\typ_i$ and identified with
  label $\lbl_i$. Following the definition in \cite{canetti03} (where the
  adversary chooses the public keys, equivalent to our $\val_i$ values), the
  functionality asks \Sim to provide the $\val_i$ values, as well as the
  identifier to associate to the complete set. If the identifier does not
  already exist in the PKI list \VDRL, the functionality fetches the
  current time from the global clock \IdealGclock, and adds the requested
  operation to the list of unconfirmed operations, \VDROP.
\item[Read.] Upon recieving the instruction \uccmd{Read}, the functionality
  fetches the dictionary associated to identifier \VDRid from the list of
  confirmed operations in the PKI. In addition, the functionality fetches the
  current time from the global clock \IdealGclock and, if there is any
  unconfirmed operation older than the predefined delay, it aborts. Otherwise,
  returns the dictionary associated to \VDRid.
\item[Update.] Receives the instruction \uccmd{Update}, associated to an
  identifier \VDRid, to which the caller whishes to add a set of label and type
  entries $\sval_a$, and from which it whishes to remove a set of labels
  $\sval_d$. If the entry \VDRid exists in the PKI list \VDRL, the functionality
  asks the adversary to provide the values for the new labels. Then, applies the
  additions and removals honestly, gets the current time from \IdealGclock, and
  adds the operation to the list of unconfirmed operations.
\item[Deactivate.] The functionality exposes the interface \uccmd{Deact}, for
  deactivating specific labels within an identifier in the PKI list \VDRL (where
  deactivating all the labels wichin an identifier is equivalent to deactivating
  the whole identifier). If the given identifier exists, the functionality
  removes the given labels from it, fetches the current time, and adds the
  operation to the list of unconfirmed operations.
\end{description}

In addition, the adversary \Sim gets access to an interface \uccmd{Process},
which lets it decide when the operations received by the ideal functionality
get applied. However, if, upon reading the contents of the PKI (via the
\uccmd{Read} interface), some unconfirmed operation is detected that is older
than the predefined delay, the functionality aborts. Note that, as long as the
maximum delay or the policy \P are not violated, the adversary has the
capability to pick in which order are (unconfirmed) operations applied, as it
specifies the operations to be processed.

\todo{The following is a bit messy, but important. Improve writeup.}
Concerning the policy \P, the functionality checks that every operation to be
processed satisfies it. Policies (may) take into account both the contents of
the operation ($\VDROp = (\VDRid,\sval,P,\tau)$) and the current (confirmed)
state of the PKI, \VDRL. Note that the specific set of values \sval in the
unconfirmed operation are always computed in advance by the corresponding
interface (\uccmd{Create}, \uccmd{Update}, or \uccmd{Deactivate}), ensuring that
the value passed to the policy has been correctly processed. In addition, the
functionality does not make any check on the identity of the party who submitted
each operation. Instead, this can be done by the policies, and allows capturing
multiple types of PKIs, e.g., PKIs that support delegation, or those that don't.
\todo{We should at least specify a bit more formally what conditions policies
  need to meet in this regard? How?}

\begin{figure}
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{Upon receiving a call $\IdealGPKIDID(\cmd,\sid,\arg)$
          from some party $P$ to run command \cmd, in session \sid, with arguments
          \arg, $\IdealGPKIDID$ operates as follows.}
        \textrm{If the received \cmd and \arg pair does not meet any of
          the following options, the request is ignored.}
        \textrm{\P is a boolean predicate defining concrete conditions for
          acceptable operations.}
        \textrm{\didOP is a list of submitted (yet, possibly unconfirmed)
          operations, and \VDRL is the authoritative state of published
          operations.}        
      \end{minipage}
    }
    \vspace*{0.5em}

    \scalebox{0.9}{
      \begin{minipage}[t]{0.57\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \setind{(\lbl_i,\typ_i)}{i\in[n]}$}{ 
          \ucsend~(\uccmd{Create}, \sid, \arg)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{CreateOk}, \sid, \VDRid,
          \sval=\setind{\val_i}{i\in[n]}) \\
          \hspace*{2.7em}\text{from}~\Sim \\
          \pcif (\VDRid,\cdot,\cdot,\cdot) \in \VDRL: \pcreturn \bot \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\
          \VDROP \gets \VDROP \cup \set{(\VDRid,
            \setind{(\lbl_i,\typ_i,\val_i)}{i\in[n]},P,\tau)} \\
          \ucio{Output}~(\uccmd{ProcessedC},\sid,\VDRid,\sval)
        }
        \vspace*{0.15em} \\

        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Update} \land
          \arg = (\VDRid, \sval_d,\sval_a)$}{
          \pcif (\VDRid,\sval_0,P,\cdot) \notin \VDRL: \pcreturn \bot \\
          \ucsend~(\uccmd{Update}, \sid,\VDRid,\sval_d,\sval_a)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{UpdOk},\sid,\VDRid,\setind{\val^a_i}{i\in[n^a]})~
          \text{from}~\Sim \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\
          \lbrace (\lbl_i^a,\typ_i^a) \rbrace_{i\in[n^a]} \gets \sval_a \\
          \sval_1 \gets (\sval_0 \setminus \sval_d) \cup
          \setind{(\lbl^a_i,\typ^a_i,\val^a_i)}{i\in[n^a]} \\
          \VDROP \gets \VDROP \cup \set{(\VDRid,\sval_1,P,\tau)} \\
          \ucio{Output}~(\uccmd{ProcessedU},\sid,\VDRid,\sval_1)
        }        
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Read} \land
          \arg = \VDRid$}{
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\
          \pcif \exists (\cdot,\cdot,\cdot,\tau') \in \VDROP~\suchthat~
          |\tau-\tau'| > \delta: \\
          \pcind \pcreturn \bot \\
          \pcif (\VDRid,\sval,\cdot,\cdot) \notin \VDRL: \pcreturn \bot \\
          \ucio{Output}~(\uccmd{ReadOk},\sid,\VDRid,\sval)~\text{to}~P
        }
        \vspace*{0.15em} \\

        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Deact} \land \arg =
          (\VDRid, \lbrace \lbl_i \rbrace_{i\in[n]})$}{
          \pcif (\VDRid,\sval_0,P,\cdot) \notin \VDRL: \pcreturn \bot \\ 
          \sval_1 \gets \sval_0 \setminus \lbrace \lbl_i \rbrace_{i\in[n]} \\
          \ucsend~(\uccmd{Deact}, \sid, \arg)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{DeactOk},\sid)~\text{from}~\Sim \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\
          \VDROP \gets \VDROP \cup \set{(\VDRid,\sval_1,P,\tau)} \\
          \ucio{Output}~(\uccmd{ProcessedD},\sid,\VDRid,\sval_1)
        }
        \vspace*{0.15em} \\

        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Process} \land \arg =
          \VDROP'$}{
          \pcif P \neq \Sim \lor \VDROP' \not\subseteq \VDROP: \pcreturn \bot \\
          \pcfor \VDROp = (\VDRid,\sval,P,\tau) \in \VDROP': \\
          \pcind \pcif \P(\VDROp,\VDRL) = 0: \pcreturn \bot \\
          \VDROP \gets \VDROP \setminus \VDROP' \\
          \VDRL \gets \VDRL \cup \VDROP'
        }
      \end{minipage}
    }
  \end{framed}
  \caption{Ideal PKI functionality with delayed verifiable updates,
    \IdealGPKIDID. We abuse set notation, and write $\lbrace (a_i,\dots)
    \rbrace_{i\in[n]} \setminus \lbrace (b_j,\dots) \rbrace_{j\in[m]}$ to mean
    subtracting from the first set the elements present in the second set,
    whenever $a_i = b_j$ for some $i,j$, even if the size of the $(a_i,\dots)$
    and $(b_j,\dots)$ tuples do not match.}
  \label{fig:fpkidid}
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "prism-protocol"
%%% End: