### 2022-12-19 prism-backend/research/prism-protocol/did_atala.tex
Beginning proof of Atala DIDs with global subroutines.

\section{Atala PRISM's \RealPKIDIDAtala: A DID-based PKI}
\label{sec:did-atala}

In this section we first describe Atala PRISM DID method \RealPKIDIDAtala, and
prove that it realizes \IdealGPKIDID.
%
We begin by describing the types of keys supported in Atala PRISM, and the
policy \PAtala that all CRUD operations need to meet.

\paragraph{Custom Notation.} %
Keys in Atala are identified by labels. In the following description of the
\RealPKIDIDAtala construction, given a set of keys, we frequently refer to the
``master key with lowest label'' (using any order relation, like lexicographic
order).  For brevity, given a set \st, we use $\LMKL(\st)$ to denote such label.

\paragraph{Key Types.} %
All operations in Atala PRISM are authenticated, meaning that a cryptographic
key needs to be involved in proving that the operation comes from the intended
party. Atala specifies four key types, that are thus supported in Atala's DID
method. Namely: master keys (\MasterKey), for operating on DIDs; issuing keys,
for issuing VCs (\IssueKey); communication keys (\CommKey), for key-exchange
protocols; and authentication keys, to associate to VCs (\AuthKey).

\paragraph{\PAtala policy.} %
Before proceeding with the specififying the DID operations in Atala, we define
its policy on the keys' labels and types.
\todo{This needs to take into account the caller, and the previous state of the
  PKI list.}

\begin{figure}[ht]
  \centering
  \procedure{$\PAtala(\lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]})$}{    
    \pcif \exists i\in[n]~\suchthat~\typ_i \notin [\MasterKey,\IssueKey,
    \CommKey,\AuthKey]: \pcreturn 0~\pccomment{All types must be known} \\   
    \pcif \not\exists i\in[n]~\suchthat~\typ_i = \MasterKey: \pcreturn 0~
    \pccomment{There must be at least one master key} \\
    \pcif \exists i, j \in[n], i \neq j~\suchthat~\lbl_i = \lbl_j: \pcreturn 0~
    \pccomment{No duplicate labels} \\
    \pcreturn 1
  }
  \caption{Policy \PAtala that all DID documents must meet in Atala.}
\end{figure}

\paragraph{DID Operations.} %
\todo{For now, this ignores \emph{unpublished} DIDs.}
\figref{fig:atalapkidid1} and \figref{fig:atalapkidid2} contain the
specification of Atala PRISM's DID method,
which we describe next. Looking at the \texttt{Create} and \texttt{Update}
operations, it is clear that $\P(\lbrace (i, \lbl_i) \rbrace_{i\in[n]})$ in
Atala requires that all key labels are of a known type (\MasterKey, \IssueKey,
\CommKey, or \AuthKey) and that at least one of the keys in the DID Document
is of type \MasterKey.

\begin{description}
\item{\uccmd{Create}.} %
  To create a DID, the user has to specify the fresh identifiers (labels) and
  types of the key pairs to include in the DID. The \uccmd{Create} algorithm
  creates as many keys of each type as requested, by leveraging the ideal
  functionality \IdealFSig, and keeps a list with the identifier, type and
  public key of each key pair. This list is then hashed into a value \texttt{h}
  via \IdealGRO, and the created DID is ``\texttt{\did:prism:h}''. A transaction
  containing the created list, the DID, and the \texttt{CreateDID} instruction
  is sent to the blockchain via a call to \IdealGdledger, signed by one of the
  master keys -- w.l.o.g., we 
  simply use the first one in the list. The DID owner stores locally the list.
\item[\uccmd{Read}.] %
  Getting the contents of a DID simply requires processing the associated
  transactions in the blockchain. Namely, the interface first fetches all the
  contents from the \IdealGdledger functionality. Then, transactions the
  relevant transactions are those with $(\tx,\sig)$ where $\tx$
  is either $(\texttt{CreateDID},\did,\cdot)$ or $(\texttt{UpdateDID},\did,
  \cdot)$. The associated DID Document is the result of applying in an ordered
  manner (from oldest to newest) all such validly signed transactions.
\item[\uccmd{Update}.] %
  To update a DID, we first assume for ease of exposition that the calling party
  $P$ has an updated version of the associated DID Document (which included
  a list of $n$ keys) -- if this is not the case, it suffices to make first a
  call to \uccmd{Read} for the given DID.
  %
  Then, updating a DID requires to (optionally) specify a list of previously
  existing key labels to remove $\sval_d$, and (optionally) specify a list
  $\sval_a$ of new key labels and their types to add. The result of removing
  the labels in $\sval_d$ and adding the new keys in $\sval_a$ must satisfy
  \PAtala. Key pairs for the new keys are generated via \IdealFSig, and a
  transaction is sent to the blockchain via \IdealGdledger, signed (again, via
  \IdealFSig) with one of the master keys in the DID \emph{before applying the
    update}. This transaction contains a list of the identifiers of the keys to
  remove, and a list of the identifiers and types of the keys to add.
  %
  Note that the DID remains the same, even though the original contents of
  the associated DID document may change.
\item[\uccmd{Deactivate}.] %
  A deactivate operation is simply an update operation with no new key
  identifiers, and that only specifies the existing identifiers that will
  \emph{not} be deactivated.
\end{description}

\begin{figure}[ht!]
  \begin{framed}

    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{Party $P$ running operation \cmd with arguments \arg, in
          session \sid.}
      \end{minipage}
    }
    \vspace*{0.5em}

    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]}$}{
          \pcif \PAtala(\lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]}) = 0: abort \\
          \st \gets \emptyset \\
          \pcfor i \in [n]: \\
          \pcind \sid' \gets ((P,\lbl_i),\sid) \\
          \pcind \ucsend~(\uccmd{KeyGen},\sid')~\text{to}~\IdealFSig \\
          \pcind \ucrecv~(\uccmd{VerKey},\sid',\pk_i)~\text{from}~
          \IdealFSig \\
          \pcind \st \gets \st \cup \lbrace (\lbl_i,\typ_i,\pk_i) \rbrace \\
          \ucsend~(\uccmd{Hash},\st)~\text{to}~\IdealGRO \\
          \ucrecv~(\uccmd{Hashed},\st,h)~\text{from}~\IdealGRO \\
          \did \gets ``did:prism:''||~h \\
          \tx \gets (\texttt{CreateDID}, \did, \st) \\
          \lbl \gets \LMKL(\st); \sid' \gets ((P,\lbl),\sid) \\
          \ucsend~(\uccmd{Sign},\sid',\tx)~\text{to}~\IdealFSig \\
          \ucrecv~(\uccmd{Signed},\sid',\tx,\sig)~\text{from}~\IdealFSig \\
          \ucsend~(\uccmd{Submit},(\tx,\sig))~\text{to}~\IdealGdledger \\
          \text{Locally store}~
          (\did, \lbrace (\lbl_i,\typ_i,\pk_i) \rbrace_{i\in[n]}) \\
          \ucio{Output}~(\uccmd{ProcessedC},\sid,\did,\lbrace \pk_i
          \rbrace_{i\in[n]})
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Read} \land
          \arg = did$}{
          \ucsend~(\uccmd{Retrieve})~\texttt{to}~\IdealGdledger \\
          \ucrecv~(\uccmd{Retrieve},L)~\texttt{from}~\IdealGdledger \\
          \st \gets \emptyset \\
          L_{\did} \gets (\ast,\did,\ast) \subseteq \DID \\
          \pcfor (\tx=(\texttt{op},\did,\cdot),\sig) \in L_{\did}: \\
          \pcind \pk \gets \emptyset \\
          \pcind \pcif \texttt{op} = \texttt{CreateDID}~\land \\
          \pcind \pcind \text{no \texttt{CreateDID} appeared before}: \\
          \pcind \pcind (\cdot,\cdot,\lbrace (\lbl_i,\typ_i,\pk_i)
          \rbrace_{i\in[n]}) \gets \tx \\
          \pcind \pcind \st \gets \lbrace (\lbl_i,\typ_i,\pk_i) \rbrace_{i \in [n]} \\
          \pcind \pcif \texttt{op} = \texttt{UpdateDID}: \\
          \pcind \pcind (\cdot,\cdot,(\st_d,\st_a)) \gets \tx \\
          \pcind \pcind \lbrace \lbl^d_i \rbrace_{i\in[n^d]} \gets \st_d \\ 
          \pcind \pcind \lbrace (\lbl^a_i,\typ^a_i,\pk^a_i)
          \rbrace_{i\in[n^a]} \gets \st_a \\
          \pcind \pcind \st \gets (\st \setminus \st_d) \cup \st_a \\
          \pcind \lbl \gets \LMKL(\st) \\
          \pcind \pcif \lbl = \emptyset \lor \PAtala(\st) = 0: \text{skip} \\
          \pcind \pcelse \sid' \gets ((P,\lbl),\sid) \\
          \pcind \ucsend~(\uccmd{Verif},\sid',\tx,\sig,\pk)~\text{to}~\IdealFSig \\
          \pcind \ucrecv~(\uccmd{Verifd},\sid',\tx,b)~\text{from}~\IdealFSig \\
          \pcind \pcif b = 0: \text{skip} \\
          \ucio{Output}~(\uccmd{ReadOk},\sid,\did,\st)
        }    
      \end{minipage}
    }
  \end{framed}
  \caption{\texttt{Create} and \texttt{Read} operations in Atala
    PRISM DID-based PKI protocol, \RealPKIDIDAtala, realising \IdealGPKIDID
    in the $(\IdealGRO,\IdealFSig,\IdealGdledger)$-hybrid model.}
  \label{fig:atalapkidid1}
\end{figure}

\begin{figure}[ht!]
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{Party $P$ running operation \cmd with arguments \arg, in
          session \sid.} \\
      \end{minipage}
    }
    \vspace*{0.5em}

    \scalebox{0.90}{          
      \begin{minipage}[t]{0.6\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Update} \land
          \arg = (did, \sval_d, \sval_a)$}{
          \text{Fetch}~
          (\did, \sval=\lbrace (\lbl_i,\typ_i,\pk_i) \rbrace_{i\in[n]})
          ~\text{from storage} \\
          \st_a \gets \emptyset;~
          \lbrace (\lbl_i,\typ_i,\pk_i) \rbrace_{i\in[n]} \gets \sval \\
          \lbrace \lbl^d_i \rbrace_{i\in[n_d]} \gets \sval_d;~
          \lbrace (\lbl^a_i,\typ^a_i) \rbrace_{i\in[n_a]} \gets \sval_a \\
          \sval' \gets (\sval \setminus \sval_d) \cup \sval_a
          \pccomment{\todo{Syntax refinement needed here.}} \\
          \pcif \PAtala(\sval') = 0: abort \\          
          \pcfor i \in [n_a] \\
          \pcind \ucsend~(\uccmd{KeyGen},((P,\lbl_i),\sid))~\text{to}~\IdealFSig \\
          \pcind \ucrecv~(\uccmd{VerKey},((P,\lbl_i),\sid),\pk^a_i)~\text{from}~
          \IdealFSig \\
          \pcind \st_a \gets \st_a \cup \lbrace (\lbl^a_i,\typ^a_i,\pk^a_i) \rbrace \\
          \tx \gets (\texttt{UpdateDID}, \did, \sval_d, \st_a) \\
          \lbl \gets \LMKL(\sval) \\
          \ucsend~(\uccmd{Sign},((P,\lbl),\sid),\tx)~\text{to}~\IdealFSig \\
          \ucrecv~(\uccmd{Signed},((P,\lbl),\sid),\tx,\sig)~\text{from}~\IdealFSig \\
          \ucsend~(\uccmd{Submit},(\tx,\sig))~\text{to}~\IdealGdledger \\
          \text{Locally store}~(\did,\sval) \\
          \ucio{Output}~(\uccmd{ProcessedU},\sid,\did,\lbrace \pk^a_i \rbrace_{i\in[n^a]}) \\
          \text{\todo{The previous must return all the final pks, not only
              $pk^a$, to be compatible with deact.}}
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.45\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Deact} \land \arg =
          (did, \sval)$}{       
          \text{Run as an \texttt{Update} operation with} \\
          \text{ $\sval_d \gets \sval$ and $\sval_a \gets \emptyset$,
            and output} \\
          \text{\uccmd{ProcessedD} instead of \uccmd{ProcessedU}}
        }
      \end{minipage}
    }
  \end{framed}
  \caption{\texttt{Update} and \texttt{Deactivate} operations in Atala
    PRISM DID-based PKI protocol, \RealPKIDIDAtala, realising \IdealGPKIDID
    in the $(\IdealGRO,\IdealFSig,\IdealGdledger)$-hybrid model.}
  \label{fig:atalapkidid2}
\end{figure}

\subsubsection{Notes for Implementations.} %

We acknowledge that, in the pseudocode given in \figref{fig:atalapkidid1} and
\figref{fig:atalapkidid2}, we have favoured notational simplicity over
implementation efficiency or
convenience. For instance, it is not really needed that the master key signing a
transaction is the one with the lowest index -- any such key can actually do it.
However, it makes notation much easier. Similarly, in practice it may be easier
to use different ``function arguments'' to denote the set of keys to remove and
a set of keys to add from/to an existing DID. However, notationally it is
simpler to follow the chosen (equivalent) approach. Finally, key indexes don't
need to follow a sequential order.

Somehow related, it is of course not needed to fetch the contents of the whole
blockchain whenever we want to get an up to date DID Document. Instead, the
party running the operation can keep a local copy, and just fetch the new
operations since it last synced. However, again the given simplified notion of
a blockchain (via \IdealGdledger) eases notation -- and seems good enough for a
first modelling effort.

\subsection{Security of \RealPKIDIDAtala}
\label{ssec:sec-didatala}

We prove security of \RealPKIDIDAtala by showing that it UC-emulates
\IdealGPKIDID. For this, we leverage concretely the results in \cite{bch+20} and
\cite{bhz21}. In a nutshell, the former extends the notions of UC emulation and
composition so that they are compatible, in the plain UC model, with global
subroutines (such as a global clock, a ledger, or a random oracle).
\cite{bhz21} builds on these results to make them suitable for systems that rely
simultaneously on multiple global functionalities.
%
To make the previous a bit more precise, we replicate here the relevant theorem
from \cite{bhz21} (in \cite{bch+20}, the same statement appears with only one
global subroutine).

\begin{theorem}[UC emulation with multiple global setups \cite{bhz21}]
  \label{thm:uc-bhz21}
  Let $\pi$, $\phi$ and $\gamma_1,\dots,\gamma_n$ be protocols. We say that
  $\pi$ $\xi$-UC-emulates $\phi$ in the presence of global subroutines
  $\gamma_1,\dots,\gamma_n$ if protocols $\pi$ and $\phi$ are formulated with
  respect to a global subroutine $\hat{\gamma}_{\gamma_1,\dots,\gamma_n}$ and
  $\UCM[\pi,\hat{\gamma}_{\gamma_1,\dots,\gamma_n}]$ $\xi$-emulates protocol
  $\UCM[\phi,\hat{\gamma}_{\gamma_1,\dots,\gamma_n}]$.
\end{theorem}

At a high level, to apply the results in \cite{bch+20,bzh21}, we show the global
subroutines used in \IdealGPKIDID and \RealPKIDIDAtala can be merged into the
single ``meta'' global subroutine $\hat{\gamma}_{\gamma_1,\dots,\gamma_n}$
required by \thmref{thm:uc-bhz21}, which furthermore meets the conditions to be
subsequently suitable for UC-emulation. In addition, we argue that this is
usually the case for the known realisations of each respective global
subroutine. (This ensures that our approach is natural, and indeed UC-realisable
in practice.)
%
We then describe how the management protocol \UCM works in conjunction with
our \IdealGPKIDID and the ``meta'' global subroutines (in the ideal world), and
with \RealPKIDIDAtala and the ``meta'' global subroutine (in the real world).
%
Finally, we show that the latter emulates the former.
%
We emphasise that this is merely an application of the results in
\cite{bch+20,bhz21} to our setting. The previously sketched process is detailed
next, directly referencing the relevant theorems of the relevant works, when
appropriate.

\paragraph{The meta-global subroutine \gAtala.} %
We next show that the global subroutines used in \RealPKIDIDAtala are leveraged
in a way that is compatible with \cite[Theorem 4.3]{bhz21}. This is not a full
proof for a concrete application of such theorem, as that would require proving
that concrete protocols UC-emulate each global functionality. However, we
initially leave it ``incomplete'' to show that any combination of instantiations
of our global subroutines can be used to build a secure variant of the Atala
DID protocol (later, we show the concrete instantiations that Atala currently
uses). For concreteness, we refer to this ``partial'' property as mergability
of the global subroutines of Atala into \gAtala.

\begin{lemma}[Mergability of \gAtala]
  \label{lma:gAtala}
  Let $\gamma_1 \gets \IdealGdledger$, $\gamma_2 \gets \IdealGRO$, $\gamma_3
  \gets \IdealGclock$. Then:
  \begin{enumerate}
  \item For $i < 3$, $\gamma_i$ is subroutine respecting except for calls to
    $\gamma_{i+1,\dots,3}$, and $\gamma_3$ is subroutine respecing. All
  $\gamma_i$'s are subroutine exposing.
\item For $i>1$, $\gamma_i$ is $\gamma_j$-regular, for all $j \in \set{1,
    \dots,\i-1}$.
  \end{enumerate}
\end{lemma}

\begin{proof}
  \qed
\end{proof}

% Before going into the proofs, observe that the \uccmd{Create} (in lines 1-2) and
% \uccmd{Update} (in lines 6-7) operations in \RealPKIDIDAtala include a policy on
% the received key labels and types. Concretely, this policy requires that all
% labels are unique, all types are of a known key type, and that at least one of
% them is of type \MasterKey. In the sequel, we use \PAtala to refer to this
% policy.

% \todo{Do we mention somewhere that {\sid}s have structure $(P,\sid')$? This is
%   used for \Sim to know which party is generating which keys.}

% \begin{theorem}[UC-security of \RealPKIDIDAtala]
%   \label{thm:sec-didatala}
%   \RealPKIDIDAtala UC-realizes $\IdealGPKIDID$ in the $(\IdealGRO,
%   \IdealFSig, \IdealGdledger)$-hybrid model.
% \end{theorem}

% \begin{proof}
%   We show that \RealPKIDIDAtala UC-realizes \IdealGPKIDID, when configured
%   with policy \PAtala. For this, we give a simulator \Sim such that, for
%   any environment \Env, $\Exec_{\RealPKIDIDAtala\adv,\Env} \approx
%   \Exec_{\phiDID,\Sim,\Env}$,
%   where \phiDID is a protocol composed with the same parties as
%   \RealPKIDIDAtala but where each party is simply a dummy party that forwards
%   every message from \Env to the ideal functionality \IdealGPKIDID, and every
%   response from \IdealGPKIDID back to \Env. The simulator \Sim runs internally a
%   copy of the real-world adversary \adv, as well as the ideal functionalities
%   \IdealGRO, \IdealFSig and \IdealGdledger, and any involved party.

%   \begin{description}
%   \item[Simulating DID Create.] %
%     % 
%     % Keys generation
%     If \Sim receives a message $(\uccmd{Create},\sid,\lbrace (\lbl_i,\typ_i)
%     \rbrace_{i\in[n]})$ from \IdealGPKIDID, with $\sid=(P,\sid')$, it simulates
%     the key generation process of $n$ signing key pairs run in the real
%     protocol, by sending $n$ queries $(\uccmd{KeyGen},((P,\lbl_i),\sid))$ to
%     \adv~on behalf of \IdealFSig, and receiving the $n$ corresponding
%     $(\uccmd{VerKey},((P,\lbl_i),\sid),\pk_i)$ responses.
%     %
%     % Hash computation
%     \Sim then concatenates all $(\lbl_i,\typ_i,\pk_i)$ tuples into \st, and
%     simulates a query to \IdealGRO, to set \did as ``$did:prism: h$'' as DID
%     value.
%     %
%     % Tx sim
%     With the DID value, \Sim simulates for \adv~the preparation of transaction
%     \tx. That is, \Sim uses the master key with lowest label, $\lbl \gets
%     \LMKL(\st)$, and simulates a query $(\uccmd{Sign},((P,\lbl),\sid),\tx)$ to
%     \IdealFSig, obtaining from \adv~$(\uccmd{Signed},((P,\lbl),\sid),\tx,\sig)$.
%     %
%     % BB append
%     \Sim also simulates appending the $(\tx,\sig)$ tuple to the bulletin board,
%     by sending $(\uccmd{Submit},(\tx,\sig))$ to \adv~on behalf of \IdealGdledger.
%     %
%     % Output
%     Finally, \Sim sends the $(\uccmd{CreateOk},\sid,\did,\lbrace \pk_i
%     \rbrace_{i\in[n]})$ to \IdealGPKIDID, and stores the tuple $(P,\did,\lbrace
%     (\lbl_i,\typ_i,\pk_i)\rbrace_{i\in[n]})$.
%     %
%     % Additional considerations
%     \todo{The following should probably go to the argumentation part.}
%     Note that \did is generated by an ideal random oracle (thus, uniformly at
%     random), just as in the real protocol \RealPKIDIDAtala. Hence, the
%     probability that \IdealGPKIDID halts because $\did \in \DID$ is negligible,
%     and \IdealGPKIDID outputs $(\uccmd{Created},\sid,\did,\lbrace \val_i=\pk_i
%     \rbrace_{i\in[n]})$ to \IdealGPKIDID.
%   \item[Simulating DID Read.]
%     To simulate read operations, when \Sim receives $(\uccmd{Read},\sid,\did)$
%     from \IdealGPKIDID, it first simulates for \adv~the call to \IdealGdledger, to
%     get the current status $L$ of the bulletin board. Then, \Sim reproduces the
%     processing a normal party would do: from older to newer entry in $L$ related
%     to \did, applies the updates in a sequential order -- ignoring any
%     \uccmd{UpdateDID} operation before a \uccmd{CreateDID}, and only applying
%     the first \uccmd{CreateDID}. If, after applying an update, \PAtala is
%     not met by the resulting set of keys and labels, \Sim skips the update.
%     Otherwise, \Sim simulates for \adv~a verification query to \IdealFSig for
%     the transaction corresponding to the update, and skips the update if the
%     signature fails.
%     %
%     Finally, if after processing all transactions related to \did, the resulting
%     contents differ from those stored by \Sim in its record $(P,\did,\sval)$,
%     \Sim aborts \todo{Do we need this?}. Otherwise, \Sim returns the processed
%     contents.    
%   \item[Simulating DID Update.]
%     % Load DID
%     Upon receiving a request from dummy party $P$, \IdealGPKIDID fetches
%     $(P,\did,\sval)$ from its records, and checks if $(\sval \setminus \sval_d)
%     \cup \sval_a$ meets \PAtala. If either step fails, \IdealGPKIDID
%     aborts. Else, it sends $(\uccmd{Update},\sid,\did,\sval_d,\sval_a)$ to \Sim.
%     \Sim then retrieves $(P,\did,\sval)$ from its records \todo{this is kept at
%       some point, no?)} and simulates for \adv~the calls to \IdealFSig for the
%     keys in the received $\sval_a$. \Sim gets the key with the lowest label from
%     \sval, with label $\lbl$, bundles the transaction \tx, and simulates a call
%     to \IdealFSig to get a signature over \tx. Finally, \Sim simulates the call
%     to \IdealGdledger, and returns $(\uccmd{Updated},\sid,(\did,\lbrace \val_i
%     \rbrace_{i\in[n_a]}))$ to \IdealGPKIDID, where the $\val_i$'s are the values
%     output by \adv~when creating the new keys via simulated calls to \IdealFSig.
%   \item[Simulating DID Deactivate.] Directly follows from DID Update.      
%   \end{description}
%   \todo{This to take into account when arguing indistinguishability}:
%   \begin{itemize}
%   \item The records of \Sim are always consistent with \IdealGdledger.
%   \end{itemize}

%   % General comments
%   Now, observe that, in all ideal operations, the policy \PAtala is checked
%   on the resulting data structure. If the policy is not satisfied, the
%   functionality aborts. This is also the case for the real world.

%   % Create comments
%   Regarding create operations, note that, both in the simulation and in the real
%   world, the resulting \did value is produced in a consistent way -- via a call
%   to \IdealGRO, using the contents of the DID document to create.
%   % Key generation and signature comments
%   Similarly, the (verification) keys used and signatures are defined by \adv~
%   both the real and ideal world -- in the latter, as part of \Sim's simulation
%   of \IdealFSig.

% \end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "prism-protocol"
%%% End: